<!--
NOTE (repository version):
- Personal identifiers and machine-specific absolute paths have been removed.
- Please set `PROJECT_ROOT` / input paths to point to your local data folder (e.g., `data/`).
-->

---
title: "Analysis Figure 2: Connectomic signature and spatial transcriptomics"
output: pdf_document
date: "2025-12-08"
editor_options: 
  chunk_output_type: console
---

# Load spatial/GBMap-single cell/TCGA:
Load the required packages
```{r}
reticulate::use_condaenv("Single_cell")
## Load Library
library(SPATA2)
library(Seurat)
library(reticulate)
library(tidyverse)
library(igraph)
library(kableExtra)
library(igraph)
library(readxl)

root = "<USER_HOME>/Desktop/OP-Planung/Analysis/RNA_seq_all"
top_connected <- readRDS(paste0(root, "/top_connected.RDS"))

## Import data
source("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/import.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/Deconvolution.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/Converter.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/utils.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/plot.R")

#TCGA <-  importTCGA()
Spatial = importSpatial()
#scRNA = importGBMap()

## Import colors
pycolors <- importPyColors()

## Load source files (R functions)
#source("<USER_HOME>/Library/CloudStorage/<CONTACT_EMAIL>/Meine Ablage/Tutorials/NeuroPhysiology/mea_functions.R")
#source("<USER_HOME>/Library/CloudStorage/<CONTACT_EMAIL>/Meine Ablage/Tutorials/NeuroPhysiology/MEA_run.R")
```

## ------------------------------ Single Cell Load Omars_single cell and plot connectivity signature:
```{python}
import scanpy as sc
#import tangram as tg
#import anndata as ad
#import squidpy as sq
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import matplotlib
import matplotlib.pyplot as plt
matplotlib.use("pdf")  
from matplotlib import rcParams
rcParams["font.family"] = "Helvetica"
rcParams["pdf.fonttype"] = 42  

def saveRasterized(fig, path, dpi=300):
  # Get the current figure
  # Rasterize all scatter plot artists
  for ax in fig.axes:
      for artist in ax.collections:
          artist.set_rasterized(True)

  # Save as vector PDF with raster scatter
  fig.savefig(path, dpi=dpi, bbox_inches="tight")

import sys
sys.path.append("<USER_HOME>/Desktop/Cooperations/Adam_Sonication/Adam_RProject/")
#from helper_Graph import *
#from helper_R_L import *
from helper_XeniumAnalysis import *
figures = "<USER_HOME>/Desktop/OP-Planung/Analysis/RNA_seq_all/Figures"
adata_gb = sc.read_h5ad("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_space_snRNA_processed.h5ad")

count_categories(adata_gb, "annotation_coarse")

connectivity_sig = np.asarray(r.top_connected.index)

sc.tl.score_genes(
    adata_gb,
    gene_list=connectivity_sig,
    score_name="connectivity_score",
    use_raw=True  # if raw contains normalized data
)

ax = sc.pl.umap(
    adata_gb,
    color="connectivity_score",
    cmap="magma",  # or "viridis" / "plasma" / "inferno"
    vmin=np.percentile(adata_gb.obs["connectivity_score"], 1),
    vmax=np.percentile(adata_gb.obs["connectivity_score"], 99),
    edges=False,
    size=10,
    return_fig=True
)

saveRasterized(ax, path=f'{figures}/UMAP_connectivity_score.pdf')

import pandas as pd

df = adata_gb.obs[["connectivity_score", "annotation_coarse"]].dropna()

celltype_stats = (
    df
    .groupby("annotation_coarse")["connectivity_score"]
    .agg(["mean", "median", "std", "count"])
    .sort_values("mean", ascending=False)
)

print(celltype_stats)

import statsmodels.formula.api as smf

df_mixed = adata_gb.obs[["connectivity_score", "annotation_granular", "sample"]].dropna()

# choose a reference cell type (first factor level becomes reference)
df_mixed["annotation_granular"] = df_mixed["annotation_granular"].astype("category")

model = smf.mixedlm(
    "connectivity_score ~ C(annotation_granular)",
    df_mixed,
    groups=df_mixed["sample"]
)
res = model.fit()
print(res.summary())

coef_table = res.summary().tables[1] 
coef_sorted = coef_table.sort_values("Coef.", ascending=False)
print(coef_sorted)


### Analysis of GRIN1 KCC2 NKCC1

#NKCC1: SLC12A2
ax = sc.pl.umap(
    adata_gb,
    color="SLC12A2",
    cmap="magma",  # or "viridis" / "plasma" / "inferno"
    edges=False,
    size=10,
    vmax = 30,
    return_fig=True
)
plt.show()
saveRasterized(ax, path=f'{figures}/UMAP_connectivity_score.pdf')

neuron_types = ["Neurons (Exc)", "Neurons (Inh)"]
adata_neurons = adata_gb[adata_gb.obs["annotation_coarse"].isin(neuron_types)].copy()
del adata_neurons.uns["annotation_granular_colors"]
ax = sc.pl.umap(
    adata_neurons,
    color="SLC12A2",
    cmap="magma",  # or "viridis" / "plasma" / "inferno"
    edges=False,
    size=20,
    vmax = 10,
    return_fig=True
)
plt.show()
saveRasterized(ax, path=f'{figures}/UMAP_Neurons_NKCC1.pdf')

ax = sc.pl.umap(
    adata_neurons,
    color="annotation_granular",
    edges=False,
    size=20,
    vmax = 10,
    return_fig=True
)
plt.show()
saveRasterized(ax, path=f'{figures}/UMAP_Neurons_Celltype.pdf')




import statsmodels.formula.api as smf
adata_neurons.obs["NKCC1"] = np.array(
    adata_neurons[:, "SLC12A2"].X.toarray().flatten()
)
df_mixed = adata_neurons.obs[["NKCC1", "annotation_granular", "sample"]].dropna()
df_mixed["annotation_granular"] = df_mixed["annotation_granular"].astype("category")

# choose a reference cell type (first factor level becomes reference)
df_mixed["annotation_granular"] = df_mixed["annotation_granular"].astype("category")

model = smf.mixedlm(
    "NKCC1 ~ C(annotation_granular)",
    df_mixed,
    groups=df_mixed["sample"]
)
res = model.fit()
print(res.summary())

coef_table = res.summary().tables[1] 
coef_sorted = coef_table.sort_values("Coef.", ascending=False)
print(coef_sorted)

nk_summary = (
    adata_neurons.obs
    .groupby("annotation_granular")["NKCC1"]
    .agg(["mean", "median", "std", "count"])
    .sort_values("mean", ascending=False)
)
nk_summary


wnk_spak_signature = [
    # Core kinases
    "WNK1", "WNK2", "WNK3", "WNK4",
    
    # Downstream kinases
    "STK39",     # SPAK
    "OXSR1",     # OSR1

    # NKCC/NCC transporters (activated by SPAK/OSR1 phosphorylation)
    "SLC12A2",   # NKCC1
    "SLC12A1",   # NKCC2
    "SLC12A3",   # NCC

    # KCC transporters (inhibited by WNK–SPAK signaling)
    "SLC12A4",   # KCC1
    "SLC12A5",   # KCC2
    "SLC12A6",   # KCC3
    "SLC12A7",   # KCC4

    # Chloride homeostasis regulators
    "CLCN3", "CLCN5", "CLIC1",

    # Upstream regulators of WNK/SPAK activation
    "WNK1B",     # neuronal isoform if present
    "NEDD4L",    # regulates NCC/NKCC trafficking
    "CSNK2A1",   # CK2: required for WNK activation
    "CUL3",      # Cullin3 → WNK ubiquitination complex
    "KLHL3",     # substrate adaptor for WNK degradation
]

sc.tl.score_genes(
    adata_neurons,
    gene_list=wnk_spak_signature,
    score_name="wnk_spak_signature",
    use_raw=True  # if raw contains normalized data
)

ax = sc.pl.umap(
    adata_neurons,
    color="wnk_spak_signature",
    cmap="magma",  # or "viridis" / "plasma" / "inferno"
    vmin=np.percentile(adata_neurons.obs["wnk_spak_signature"], 1),
    vmax=np.percentile(adata_neurons.obs["wnk_spak_signature"], 99),
    edges=False,
    size=15,
    return_fig=True
)
plt.show()
saveRasterized(ax, path=f'{figures}/UMAP_Neurons_wnk_spak_signature.pdf')

df_mixed = adata_neurons.obs[["wnk_spak_signature", "annotation_granular", "sample"]].dropna()
df_mixed["annotation_granular"] = df_mixed["annotation_granular"].astype("category")

# choose a reference cell type (first factor level becomes reference)
df_mixed["annotation_granular"] = df_mixed["annotation_granular"].astype("category")

model = smf.mixedlm(
    "wnk_spak_signature ~ C(annotation_granular)",
    df_mixed,
    groups=df_mixed["sample"]
)
res = model.fit()
print(res.summary())

coef_table = res.summary().tables[1] 
coef_sorted = coef_table.sort_values("Coef.", ascending=False)
print(coef_sorted)


```

```{r}

library(dplyr)
library(stringr)
library(tibble)

clean_df_WNK <- py$coef_sorted %>%
  as.data.frame() %>%
  rownames_to_column("ID") %>%
  mutate(
    clean_label = str_replace(ID, "C\\(annotation_granular\\)\\[T\\.", ""),
    clean_label = str_replace(clean_label, "\\]", "")
  ) %>% 
  filter(!ID %in% c("Intercept", "Group Var"))

clean_df_NKCC2 <- py$coef_sorted %>%
  as.data.frame() %>%
  rownames_to_column("ID") %>%
  mutate(
    clean_label = str_replace(ID, "C\\(annotation_granular\\)\\[T\\.", ""),
    clean_label = str_replace(clean_label, "\\]", "")
  ) %>% 
  filter(!ID %in% c("Intercept", "Group Var"))


clean_df_WNK$type = "WNK"
clean_df_NKCC2$type = "NKCC2"


plot_df <- rbind(clean_df_WNK, clean_df_NKCC2)

# 1. create unique order of labels by average coefficient OR by WNK or NKCC rank
label_order <- plot_df %>%
  filter(type!="NKCC2") %>% 
  arrange(Coef.) %>%
  pull(clean_label)

# 2. apply this order as factor levels
plot_df <- plot_df %>%
  mutate(clean_label = factor(clean_label, levels = label_order))

plot_df$clean_label <- str_replace_all(plot_df$clean_label, "-", " ")

ggplot(plot_df, aes(x = clean_label, y = as.numeric(Coef.), fill = type, size = Coef.)) +
  geom_point(shape = 21 , color = "black") +
  scale_fill_manual(values = c("WNK" = "#F8766D", "NKCC2" = "#619CFF")) +
  labs(
    x = "",
    y = "Coefficient (Mixed Effects Model)",
    fill = "Signature",
    title = "WNK–SPAK and NKCC1/NKCC2 Enrichment Across Neuronal Subtypes"
  ) +
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black", angle=90, hjust=1),
          axis.text.y = element_text(colour="black"))


```




## ------------------------------ Spatial Data Analyis
```{r}
reticulate::use_condaenv("cell2loc_mps")
## Load Library
library(SPATA2)
library(Seurat)
library(reticulate)
library(tidyverse)
library(igraph)
library(kableExtra)
library(igraph)
library(readxl)

root = "<USER_HOME>/Desktop/OP-Planung/Analysis/RNA_seq_all"
top_connected <- readRDS(paste0(root, "/top_connected.RDS"))

## Import data
source("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/import.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/Deconvolution.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/Converter.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/utils.R")
#source("<USER_HOME>/Desktop/MERFISH/EcoFoundation/R_Functions/plot.R")

#TCGA <-  importTCGA()
Spatial = importSpatial()
#scRNA = importGBMap()

## Import colors
pycolors <- importPyColors()

## Load source files (R functions)
#source("<USER_HOME>/Library/CloudStorage/<CONTACT_EMAIL>/Meine Ablage/Tutorials/NeuroPhysiology/mea_functions.R")
#source("<USER_HOME>/Library/CloudStorage/<CONTACT_EMAIL>/Meine Ablage/Tutorials/NeuroPhysiology/MEA_run.R")



```


```{python}
from spatialdata import read_zarr
import scanpy as sc
import pandas as pd
import numpy as np
import squidpy as sq
import numpy as np
root = "<USER_HOME>/Desktop/Cooperations/Essen_proBcells/Analysis"
figures = "<USER_HOME>/Desktop/OP-Planung/Analysis/RNA_seq_all/Figures"

adata_vis = sq.read.visium("<USER_HOME>/Desktop/RecurrentGBM/Data/269_T/outs")
adata_vis.var_names_make_unique()
sq.pl.spatial_scatter(adata_vis, color="GFAP")
plt.show()

## Add connectivity signature

connectivity_sig = np.asarray(r.top_connected.index)

sc.tl.score_genes(
    adata_vis,
    gene_list=connectivity_sig,
    score_name="connectivity_score"
)
sq.pl.spatial_scatter(adata_vis, color="connectivity_score");plt.show()


```

## Create spatial plots:
```{python}
from cell2location.plt import plot_spatial
import os
import matplotlib
import matplotlib.pyplot as plt
matplotlib.use("pdf")  
from matplotlib import rcParams
rcParams["font.family"] = "Helvetica"
rcParams["pdf.fonttype"] = 42  

def saveRasterized(fig, path, dpi=300):
  # Get the current figure
  # Rasterize all scatter plot artists
  for ax in fig.axes:
      for artist in ax.collections:
          artist.set_rasterized(True)

  # Save as vector PDF with raster scatter
  fig.savefig(path, dpi=dpi, bbox_inches="tight")




# Directory containing the h5ad files
data_dir = "<USER_HOME>/Desktop/RecurrentGBM/Data/GBM_all"
files = [f for f in os.listdir(data_dir) if f.endswith("_adata_annotated.h5ad")]
files = sorted(files)  # optional but helpful for reproducibility

adatas = []  # to store each loaded object
IDs = []

for f in files:
    path = os.path.join(data_dir, f)
    sample = f.replace("_adata_annotated.h5ad", "")
    IDs.append(sample)
    adata = sc.read_h5ad(path)
    adata.obs["sample"] = sample
    adata.obs["file_id"] = f
    adatas.append(adata)

print(f"Loaded {len(adatas)} samples.")

i=24
print(IDs[i])
adata_vis = adatas[i]
adata_vis.obs[adata_vis.uns['mod']['factor_names']] = adata_vis.obsm["q05_cell_abundance_w_sf"]
adata_vis.obsm["q05_cell_abundance_w_sf"].columns
clust_labels = ['NPC-neuronal-like', 'Astrocytes', 'Neurons (Exc)', 'Neurons (Inh)', 'Gliosis-like', 'Hypoxic']
clust_col = ['' + str(i) for i in clust_labels]
sc.tl.score_genes(
    adata_vis,
    gene_list=connectivity_sig,
    score_name="connectivity_score"
)
sq.pl.spatial_scatter(adata_vis, color="connectivity_score", vmin = -0, vmax = 0.2);plt.show()
saveRasterized(plt.gcf(), f'{figures}/{IDs[i]}_connectivity_score.pdf')
plot_spatial(adata=adata_vis,
      color=clust_col, 
      labels=clust_labels,
      show_img=True,
      style='fast',
      max_color_quantile=0.992,
      circle_diameter=4,
      colorbar_position='right'
  )
plt.show()
saveRasterized(plt.gcf(), f'{figures}/{IDs[i]}_spatial_cells.pdf')



## Statistical testing: 


import numpy as np
import pandas as pd
from scipy.spatial import KDTree

def spatial_coloc_test_all_celltypes(
    adata,
    celltypes,
    score_key="connectivity_score",
    radius=100,
    n_perm=2000
):
    """
    Spatial co-localisation test between each cell type abundance
    and a continuous signature (e.g., connectivity score) in a Visium dataset.

    Parameters
    ----------
    adata : AnnData
        Visium AnnData object with .obsm["spatial"] and cell type abundances in .obs.
    celltypes : list
        List of obs-column names with cell-type abundance values.
    score_key : str
        obs-column with continuous signature score.
    radius : float
        Neighborhood radius (Visium spot ~55 µm → 100 µm is common).
    n_perm : int
        Number of permutations.

    Returns
    -------
    pd.DataFrame
        columns: cell_type, observed, p_value, effect_size, n_perm
    """

    coords = adata.obsm["spatial"]
    tree = KDTree(coords)
    neighbors = tree.query_ball_tree(tree, r=radius)

    # Preload connectivity values
    conn = adata.obs[score_key].values

    results = []

    for ct in celltypes:
        if ct not in adata.obs.columns:
            print(f"Warning: cell type {ct} not found; skipping.")
            continue

        abund = adata.obs[ct].values

        # -------- Observed statistic --------
        obs_vals = []
        for i in range(len(coords)):
            neigh_idx = neighbors[i]
            obs_vals.append(np.mean(abund[neigh_idx] * conn[neigh_idx]))
        observed = np.mean(obs_vals)

        # -------- Permutations --------
        perm_stats = []
        for _ in range(n_perm):
            perm_conn = np.random.permutation(conn)
            vals = []
            for i in range(len(coords)):
                neigh_idx = neighbors[i]
                vals.append(np.mean(abund[neigh_idx] * perm_conn[neigh_idx]))
            perm_stats.append(np.mean(vals))

        perm_stats = np.array(perm_stats)

        # permutation p-value (one-sided: enrichment)
        p_value = np.mean(perm_stats >= observed)

        # standardized effect size
        effect = (observed - np.mean(perm_stats)) / (np.std(perm_stats) + 1e-9)

        results.append({
            "cell_type": ct,
            "observed": observed,
            "p_value": p_value,
            "effect_size_z": effect,
            "n_perm": n_perm
        })

    return pd.DataFrame(results)


celltypes = adata_gb.obs["annotation_coarse"].values.unique()

res_df = spatial_coloc_test_all_celltypes(
    adata_vis,
    celltypes,
    score_key="connectivity_score",
    radius=100,
    n_perm=100
)

print(res_df.sort_values("p_value"))





```


## ------------------------------ Electrogenomics
```{r}
reticulate::use_condaenv("cell2loc_mps")
library(readxl)
library(patchwork)
library(tidyverse)
library(SPATA2)
library(igraph)
root_out = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/Analysis_DHH"
root_SPATA = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/data/proc"
root_MEA = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/UKA_MEA/"
meta <- readRDS("<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/Analysis_DHH/Meta.RDS")
meta[,2] <- str_replace_all(meta[,2], "dieterhenrikheiland", "<REDACTED>")
meta[,3] <- str_replace_all(meta[,3], "dieterhenrikheiland", "<REDACTED>")
meta[,5] <- str_replace_all(meta[,5], "dieterhenrikheiland", "<REDACTED>")
meta[,6] <- str_replace_all(meta[,6], "dieterhenrikheiland", "<REDACTED>")
meta[,7] <- str_replace_all(meta[,7], "dieterhenrikheiland", "<REDACTED>")
meta[,8] <- str_replace_all(meta[,8], "dieterhenrikheiland", "<REDACTED>")
meta$spata2_new <- paste0(root_out, "/", meta$samples, "_spata2.RDS")

root = "<USER_HOME>/Desktop/OP-Planung/Analysis/RNA_seq_all"
top_connected <- readRDS(paste0(root, "/top_connected.RDS"))


colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4_new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")



```

## Transfer lables and colors:
```{r}

library(dplyr)

granular_to_coarse <- c(
  # Tumor / glial-like
  "AC_like"                      = "AC-progenitor-like",
  "AC_like_Prolif"               = "Proliferative",
  "MES_like_hypoxia_independent" = "Hypoxic",
  "MES_like_hypoxia_MHC"         = "Hypoxic",
  "NPC_like_OPC"                 = "OPC-NPC-like",
  "NPC_like_Prolif"              = "Proliferative",
  "NPC_like_neural"              = "NPC-neuronal-like",
  "OPC_like"                     = "OPC-like",
  "OPC_like_Prolif"              = "Proliferative",

  # Myeloid / macrophage / DC
  "Mast"                         = "Myeloid",
  "Mono_anti_infl"               = "Myeloid",
  "Mono_hypoxia"                 = "Myeloid",
  "Mono_naive"                   = "Myeloid",
  "TAM_BDM_INF"                  = "Myeloid",
  "TAM_BDM_MHC"                  = "Myeloid",
  "TAM_BDM_anti_infl"            = "Myeloid",
  "TAM_BDM_hypoxia_MES"          = "Myeloid",
  "TAM_MG_aging_sig"             = "Myeloid",
  "TAM_MG_pro_infl_I"            = "Myeloid",
  "TAM_MG_pro_infl_II"           = "Myeloid",
  "TAM_MG_prolif"                = "Myeloid",
  "DC1"                          = "Myeloid",
  "DC2"                          = "Myeloid",
  "DC3"                          = "Myeloid",
  "cDC1"                         = "Myeloid",
  "cDC2"                         = "Myeloid",
  "pDC"                          = "Myeloid",

  # Vascular / stromal
  "Endo_arterial"                = "Vascular-associated",
  "Endo_capilar"                 = "Vascular-associated",
  "Pericyte"                     = "Vascular-associated",
  "Perivascular_fibroblast"      = "Vascular-associated",
  "SMC"                          = "Vascular-associated",
  "SMC_COL"                      = "Vascular-associated",
  "SMC_prolif"                   = "Vascular-associated",
  "Scavenging_endothelial"       = "Vascular-associated",
  "Scavenging_pericyte"          = "Vascular-associated",
  "Tip_like"                     = "Vascular-associated",
  "VLMC"                         = "Vascular-associated",

  # Canonical CNS glia
  "Astrocyte"                    = "Astrocytes",
  "OPC"                          = "OPCs",
  "Oligodendrocyte"              = "Oligodendrocytes",
  "RG"                           = "AC-progenitor-like",  # or NPC-neuronal-like if you prefer

  # T / B / NK = lymphocytes
  "NK"                           = "Lymphocytes",
  "Prolif_T"                     = "Lymphocytes",
  "Reg_T"                        = "Lymphocytes",
  "CD8_EM"                       = "Lymphocytes",
  "CD8_NK_sig"                   = "Lymphocytes",
  "CD8_cytotoxic"                = "Lymphocytes",
  "B_cell"                       = "Lymphocytes",
  "Plasma_B"                     = "Lymphocytes",
  "CD4_INF"                      = "Lymphocytes",
  "CD4_rest"                     = "Lymphocytes",

  # Stress / gliosis
  "Stress_sig"                   = "Gliosis-like",

  # Neurons – generic
  "Neuron"                       = "Neurons (Exc)",

  # Excitatory subtypes
  "L2_3_IT"                      = "Neurons (Exc)",
  "L5_ET"                        = "Neurons (Exc)",
  "L5_IT"                        = "Neurons (Exc)",
  "L5_6_NP"                      = "Neurons (Exc)",
  "L6_CT"                        = "Neurons (Exc)",
  "L6_IT"                        = "Neurons (Exc)",
  "L6b"                          = "Neurons (Exc)",

  # Inhibitory subtypes
  "Lamp5"                        = "Neurons (Inh)",
  "Pvalb"                        = "Neurons (Inh)",
  "Sncg"                         = "Neurons (Inh)",
  "Sst"                          = "Neurons (Inh)",
  "Vip"                          = "Neurons (Inh)"
)

colors

lable_exhange = granular_to_coarse %>% as.data.frame() %>% rownames_to_column("annotation_level_4_new")
names(lable_exhange)[2] <- "annotation_coarse"

cols_lab = data.frame(annotation_coarse = py$adata_gb$obs$annotation_coarse %>% levels(),
                      colors = py$adata_gb$uns$annotation_coarse_colors)
lable_exhange <- lable_exhange %>% left_join(., cols_lab)
colors_new <- left_join(data.frame(annotation_level_4_new = colors$annotation_level_4_new), lable_exhange)
colors$colors = colors_new$colors
colors$annotation_coarse = colors_new$annotation_coarse

```


## Colors
```{python}
import matplotlib.pyplot as plt
cmap = plt.get_cmap('tab20')
tab20_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab20_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab20_hex_colors

cmap = plt.get_cmap('tab10')
tab10_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab10_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab10_hex_colors
```

Samples die keinen Tumor haben !!

S24_06_CT_01_spata2
S24_06_CT_02_spata2

## Create SPATA files
```{r}
i=1
for(i in 1:nrow(meta)){
  
  raw <- readRDS(meta$raw_data[i])
  scale=1
  object <- SPATA2::initiateSpataObject_CountMtr(coords_df = raw$coords,
                                                 count_mtr=raw$counts,
                                                 image = raw$image$mea_overlay_orient,
                                                 sample_name = meta$samples[i]
                                                 )
  object <- object %>% flipImage(axis="h")
  plotSurface(object)
  
  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))
  
}
meta$spata2_new <- paste0(root_out, "/", meta$samples, "_spata2.RDS")


## Add cytospace:
for(i in 1:nrow(meta)){
  object <- readRDS(meta$spata2_new[i])
  object@spatial[[1]]$cytospace = readRDS(meta$cytospace[i])
  object@spatial[[1]]$graph = readRDS(meta$graph[i])
  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))

}

## Add c2l:
for(i in 1:nrow(meta)){
  object <- readRDS(meta$spata2_new[i])
  output <- paste0(root_out, "/", meta$samples[i] ,"_deconv.csv")
  cell_types <- read.csv(output)
  names(cell_types)[1] <- "barcodes"
  cell_types_df <- cell_types
  names(cell_types_df) <- str_remove(names(cell_types_df), "q05cell_abundance_w_sf_")
  names(cell_types_df) <- names(cell_types_df) %>% str_replace_all(., "[.]", "_")
  object <- object %>% SPATA2::addFeatures(cell_types_df, overwrite = T)

  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))

}

```


## Plot Single cell and mea data:
```{r}
#Colors:
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
i=6
object <- readRDS(meta$spata2_new[i])
plotSurface(object)

object@used_genesets = rbind(data.frame(ont = "connectivity", gene=top_connected$gene), object@used_genesets)

limits=c(0.1, 0.8)
p = plotSurface(object, color_by = "connectivity", alpha_by="connectivity", display_image=T, limits=limits, oob=scales::squish)+
  scale_alpha(limits=limits,oob=scales::squish)
p


cell_types <- object@spatial[[1]]$cytospace
s = plotSurface(object, pt_alpha=0)
s

library(ggforce)
p=#ggplot()+
  s+
  geom_voronoi_tile(data =cell_types, mapping=aes(x=x_sc,y=y_sc, group = -1L, fill = CellType), 
                      max.radius = 3, colour = 'black', linewidth=0.1, alpha=0.4)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")+
  Seurat::NoLegend()
p

p=ggplot()+
  geom_voronoi_tile(data =cell_types, mapping=aes(x=x_sc,y=y_sc, group = -1L, fill = CellType), 
                      max.radius = 3, colour = 'black', linewidth=0.1, alpha=1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")+
  Seurat::NoLegend()

p

## Add MEA Grid
library(ggnewscale)
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges

cond <- nodes$drug %>% unique()
drug_select = c(nodes$drug %>% unique())[1]
weight_lim = 10

n_inp = nodes[!is.na(nodes$drug), ] %>% filter(.$drug == drug_select)
e_inp = edges %>% filter(.$drug==drug_select) %>% filter(weight>weight_lim)

dim(e_inp)

p+
  ggnewscale::new_scale_color()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=1)+
  geom_segment(data = e_inp %>% arrange(weight), 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight), alpha=0.1)+ 
  scale_size(range=c(0.1,1.8))+
  scale_color_viridis_c(option="B", 
                        limits=c(0, 1000), 
                        oob=scales::squish)

ggplot()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=2, color="black")+
  geom_segment(data = e_inp, 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight))+ 
  scale_size(range=c(0.1,1))+
  scale_color_viridis_c(option="B", limits=c(0.3, 0.7), oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

```

## KNN analysis
```{r}
#Colors:
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
i=6
object <- readRDS(meta$spata2_new[i])
plotSurface(object)
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges

## Get KNN cells of the MEA spots
library(FNN)
drug_select = c(nodes$drug %>% unique())[3]
n_inp = 
  nodes[!is.na(nodes$drug), ] %>% 
  filter(.$drug == drug_select)

#node_rank = n_inp %>% arrange(desc(degree)) %>% mutate(degree_rank=1:nrow(.)) %>% select(id, degree_rank)
node_rank = n_inp %>% arrange(desc(nonrandom_spikes)) %>% mutate(degree_rank=1:nrow(.)) %>% select(id, degree_rank)


query_points = data.frame(x=n_inp$x, y=n_inp$y)
#data_points = data.frame(x=cell_types$x_sc, y=cell_types$y_sc)
coords = SPATA2::getCoordsDf(object)
data_points = data.frame(x=coords$x, y=coords$y)
k <- 5

# Find k-nearest neighbors of query_points in data_points
knn_result <- get.knnx(data = data_points, query = query_points, k = k)
index = knn_result$nn.index
dist = knn_result$nn.dist   


NN = map_dfr(1:nrow(index), .f=function(i){
  print(i)
  data.frame(
    barcodes = coords$barcodes[i],
    degree = n_inp$degree[index[i,]] %>% mean(na.rm=T),
    random = n_inp$nonrandom_spikes[index[i,]]%>% mean(na.rm=T),
    firing_rate = n_inp$firing_rate[index[i,]]%>% mean(na.rm=T)
  )
})

object@used_genesets = rbind(data.frame(ont = "connectivity", gene=top_connected$gene), object@used_genesets)
NN <- left_join(NN, joinWithGeneSets(object, gene_sets = "connectivity"), by="barcodes")




ggplot(NN,mapping = aes(x=connectivity, y=degree))+geom_point()+
  geom_smooth(se=F, method="lm")+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))

cor.test(NN$degree, NN$connectivity)


cell_types <- object@spatial[[1]]$cytospace
NN = map_dfr(1:nrow(index), .f=function(i){
  data.frame(
    UniqueCID = cell_types$UniqueCID[index[i,]],
    dist = dist[i,],
    id = n_inp$id[i],
    degree_rank = node_rank %>% filter(id==n_inp$id[i]) %>% pull(degree_rank) %>% mean(),
    degree = n_inp$degree[i],
    random = n_inp$nonrandom_spikes[i],
    firing_rate = n_inp$firing_rate[i]
  )
})
NN <- left_join(NN, cell_types, by="UniqueCID")
NN <- left_join(NN, data.frame(CellType = colors$annotation_level_4_new, anno2 = colors$annotation_coarse))




plot_df <- NN %>% group_by(degree_rank, anno2) %>% count()
plot_df$anno2 <- factor(plot_df$anno2, levels = unique(colors$annotation_coarse))

cc2 <- colors$colors;names(cc2) <- colors$annotation_coarse
plot_df
ggplot()+
  geom_col(plot_df, mapping=aes(x=as.factor(degree_rank), y=n, fill=anno2), position = "fill")+
  scale_fill_manual(values=cc2)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))



plot_df <- plot_df %>% 
                filter(anno2 %in% c("Neurons (Exc)", "Neurons (Inh)", "NPC-neuronal-like"))
                #filter(degree_rank>100)

ggplot()+
  geom_smooth(plot_df, 
              mapping=aes(x=degree_rank, y=n, color=anno2))+
  scale_color_manual(values=cc2)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))


plot_df <- plot_df
ggplot()+
  geom_smooth(plot_df, 
              mapping=aes(x=degree_rank, y=n, color=anno2), se=F)+
  scale_color_manual(values=cc2)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))

```


## Correlation of cell abundance and network parameter

```{r}
library(FNN)
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
getCellCor <- function(object, nodes, cell_types, k=2, drug_select, prams){

  cells <- 
    object %>% 
    joinWithFeatures(features=cell_types)
  
  # subset drug
  n_inp = 
    nodes[!is.na(nodes$drug), ] %>% 
    filter(.$drug == drug_select) %>% 
    group_by(id,x,y) %>% 
    select(id,x,y, {prams}) %>% 
    summarise_all(.funs = mean)
  
  ggplot()+
    geom_point(data = data.frame(x=cells$x, y=cells$y), mapping=aes(x,y))+
    geom_point(data = n_inp[,c("x", "y")],mapping=aes(x,y), color="red")
  
  ## KNN
  knn_result <- get.knnx(data = data.frame(x=cells$x, y=cells$y), query = n_inp[,c("x", "y")] , k = k)
  
  ## Summarize
  NN = map_dfr(1:nrow(knn_result$nn.index), .f=function(i){
      
    cell_mat = 
      cells[knn_result$nn.index[i,], cell_types] %>% colSums() %>% 
      as.data.frame()
    names(cell_mat)[1] <- "fraction"
    cell_mat <- cell_mat %>% rownames_to_column("Cell_type")
    
    cell_mat <- 
      cell_mat %>% 
      mutate(
        dist = mean(knn_result$nn.dist[i,]),
        id = n_inp$id[i])
      
    return(cell_mat)
})
  NN <- NN %>% left_join(., n_inp, by="id")
  
  
  ## run cor:
  all = map_dfc(cell_types, .f=function(ct){
    
    inp=na.omit(NN)
    cor_mat <- inp %>% 
      filter(Cell_type==ct) %>%
      select(fraction, {prams}) %>% 
      cor()
    out = cor_mat[1,] %>% as.data.frame()
    names(out)=ct
    return(out)
  })
  all <- all[-c(1), ]
  return(all)
}


i=1
for(i in 1:6){
  print(i)
  object <- readRDS(meta$spata2_new[i])
  nodes <- object@spatial[[1]]$graph$nodes
  edges <- object@spatial[[1]]$graph$edges
  
  all_drugs <- as.character(na.omit(nodes$drug %>% unique()))
  
  all_drugs_list = map(all_drugs, .f=function(drug_select){
    prams <- names(nodes)[c(18:24, 27:46)]
  cell_types <- names(cc)

  all = getCellCor(object, nodes, cell_types, k=4, drug_select, prams)

  all_sub <- 
    all %>% t() %>% 
    as.data.frame() %>% 
    rownames_to_column("annotation_level_4_new") %>% 
    left_join(., colors) %>% 
    select(-annotation_level_4_new) %>% 
    select(annotation_coarse, {prams}) %>% 
    group_by(annotation_coarse) %>% 
    summarise_all(.funs = mean) %>% 
    as.data.frame()
  })
  names(all_drugs_list) <- all_drugs
  
  saveRDS(all_drugs_list, paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))
  
  
}



matrix_list1 <- map(c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$nodrug 
  rownames(all_sub) <- all_sub$annotation_coarse
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })

matrix_list2 <- map(c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$`8mMK`
  rownames(all_sub) <- all_sub$annotation_coarse
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })
matrix_list3 <- map(c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$`5uMgabazine`
  rownames(all_sub) <- all_sub$annotation_coarse
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })

matrix_list <- matrix_list3
all_sub <- Reduce("+", matrix_list) / length(matrix_list)
all_sub <- all_sub[colors$annotation_coarse %>% unique(), ] %>% t()
mat = all_sub[c("degree_centrality", "firing_rate", "nonrandom_spikes"),] %>% t() %>% na.omit()

pheatmap::pheatmap(mat,
                   cluster_rows = T, cluster_cols = T,
                   color = colorRampPalette(RColorBrewer::brewer.pal(9, "PiYG") )(50))




```


























```{r}
reticulate::use_condaenv("Single_cell")
```



```{python}
import squidpy as sq
import scanpy as sc
import matplotlib.pyplot as plt
adata_vis = sc.read_h5ad("test.h5ad")

# compute KNN using the cell2location output stored in adata.obsm
sc.pp.neighbors(adata_vis, use_rep='q05_cell_abundance_w_sf',n_neighbors = 15)
sc.tl.leiden(adata_vis, resolution=1.1)
adata_vis.obs["region_cluster"] = adata_vis.obs["leiden"].astype("category")

sq.pl.spatial_scatter(adata_vis, color="region_cluster")
plt.show()

adata_vis.write_h5ad("test.h5ad")

```


```{r}
reticulate::use_condaenv("cell2loc_mps")


data_all = "<USER_HOME>/Desktop/RecurrentGBM/Data/GBM_all"

path_h5ad = dir(data_all, "annotated.h5ad")

```



## Niche Analysis
```{python}
import matplotlib
import matplotlib.pyplot as plt
matplotlib.use("pdf")  
from matplotlib import rcParams
rcParams["font.family"] = "Helvetica"
rcParams["pdf.fonttype"] = 42  
import scanpy as sc
import pandas as pd
import numpy as np
import squidpy as sq
import numpy as np
import cell2location
from cell2location import run_colocation
import os
import anndata as ad

run_name = '<USER_HOME>/Desktop/Cooperations/Essen_proBcells/Analysis'
adata_vis = sc.read_h5ad("test.h5ad")
mod = cell2location.models.Cell2location.load("test", adata_vis)


# Directory containing the h5ad files
data_dir = "<USER_HOME>/Desktop/RecurrentGBM/Data/GBM_all"
files = [f for f in os.listdir(data_dir) if f.endswith("_adata_annotated.h5ad")]
files = sorted(files)  # optional but helpful for reproducibility

adatas = []  # to store each loaded object

for f in files:
    path = os.path.join(data_dir, f)
    sample = f.replace("_adata_annotated.h5ad", "")
    adata = sc.read_h5ad(path)
    adata.obs["sample"] = sample
    adata.obs["file_id"] = f
    adatas.append(adata)

print(f"Loaded {len(adatas)} samples.")
adata_all = ad.concat(adatas, join="outer", label="sample_id", keys=[a.obs["sample"][0] for a in adatas])
adata_all.var_names_make_unique()



res_dict, adata_vis = run_colocation(
    adata_all,
    model_name='CoLocatedGroupsSklearnNMF',
    train_args={
      'n_fact': np.arange(3, 30), # IMPORTANT: use a wider range of the number of factors (5-30)
      'sample_name_col': 'sample', # columns in adata_vis.obs that identifies sample
      'n_restarts': 3 # number of training restarts
    },
    # the hyperparameters of NMF can be also adjusted:
    model_kwargs={'alpha': 0.01, 'init': 'random', "nmf_kwd_args": {"tol": 0.000001}},
    export_args={'path': f'{run_name}/CoLocatedComb/'}
)

# Here we plot the NMF weights (Same as saved to `cell_type_fractions_heatmap`)
res_dict['n_fact8']['mod'].plot_cell_type_loadings()
plt.show()

print(res_dict.keys())
print(adata_vis.obsm.keys())

factors_to_plot = factor_df.columns[:4].tolist()

res_dict['n_fact8']


plt.close("all")
mod.plot_history()
plt.show()


samples = mod.sample_posterior(num_samples=1000,batch_size=2000)
samples.keys()

# Compute expected expression per cell type
expected_dict = mod.module.model.compute_expected_per_cell_type(samples["post_sample_q05"], mod.adata_manager)

# Add to anndata layers
for i, n in enumerate(mod.factor_names_):
    adata_vis.layers[n] = expected_dict['mu'][i]

adata_vis.write_h5ad("test.h5ad")

ctypes = ['Pro-B cells', 'AC-gliosis-like', 'Neurons (Inh)', 'CMP','OPC-like', 'OPC-NPC-like', 'Vascular-associated']
genes = ['CD34']
adata_vis.var["SYMBOL"] = adata_vis.var.index

sq.pl.spatial_scatter(adata_vis, color="CD163", layer = "Myeloid")
plt.show()

sq.pl.spatial_scatter(adata_vis, color="RBFOX3", layer = "Neurons (Exc)")
plt.show()

sq.pl.spatial_scatter(adata_vis, color="CD19", layer = "Pro-B cells", vmax = 0.1)
plt.show()


```


```{python}
import matplotlib.pyplot as plt
import scanpy as sc
import numpy as np


def plot_genes_per_cell_type(slide, genes, ctypes):
    n_genes = len(genes)
    n_ctypes = len(ctypes)
    fig, axs = plt.subplots(
        nrows=n_genes, ncols=n_ctypes + 1, figsize=(4.5 * (n_ctypes + 1) + 2, 5 * n_genes + 1), squeeze=False
    )
    # axs = axs.reshape((n_genes, n_ctypes+1))

    # plots of every gene
    for j in range(n_genes):
        # limit color scale at 99.2% quantile of gene expression (computed across cell types)
        quantile_across_ct = np.array(
            [
                np.quantile(slide.layers[n][:, slide.var["SYMBOL"] == genes[j]].toarray(), 0.992)
                for n in slide.uns["mod"]["factor_names"]
            ]
        )
        quantile_across_ct = np.partition(quantile_across_ct.flatten(), -2)[-2]
        sc.pl.spatial(
            slide,
            cmap="magma",
            color=genes[j],
            # layer=ctypes[i],
            gene_symbols="SYMBOL",
            ncols=4,
            size=1.3,
            img_key="hires",
            # limit color scale at 99.2% quantile of gene expression
            vmin=0,
            vmax="p99.2",
            ax=axs[j, 0],
            show=False,
        )

        # plots of every cell type
        for i in range(n_ctypes):
            sc.pl.spatial(
                slide,
                cmap="magma",
                color=genes[j],
                layer=ctypes[i],
                gene_symbols="SYMBOL",
                ncols=4,
                size=1.3,
                img_key="hires",
                # limit color scale at 99.2% quantile of gene expression
                vmin=0,
                vmax=quantile_across_ct,
                ax=axs[j, i + 1],
                show=False,
            )
            axs[j, i + 1].set_title(f"{genes[j]} {ctypes[i]}")

    return fig, axs

```

Niche Analysis

```{r}
library(reticulate)
library(tidyverse)



enrichment = py$adata_all$obsm["q95_cell_abundance_w_sf"]
names(enrichment) <- names(enrichment) %>% str_remove(., "q95cell_abundance_w_sf_")
names(enrichment) <- names(enrichment) %>% str_replace_all("-", "_")
enrichment <- enrichment[, !c(names(enrichment) %in%c("Intermediate macrophages", "Tissue Res. Macrophages"))]
all_cells = names(enrichment)

mat <- cor(enrichment, method = "spearman")
#mat <- cor(enrichment)
p=pheatmap::pheatmap(mat)

mat <- mat[p$tree_row$order, p$tree_col$order]

# Keep only lower triangle (without diagonal)
#mat[upper.tri(mat)] <- NA

# Convert to long format for ggplot
library(reshape2)
melted <- melt(mat, na.rm = TRUE)
melted[melted==1]=NA
# Plot

col = viridis::plasma(50)
col = colorRampPalette(c("#FFFFFF", RColorBrewer::brewer.pal(9, "Reds")))(50)

library(ggplot2)
ggplot(melted, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  coord_equal()+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))+
    scale_fill_gradientn(colours = col, 
                        limits=c(0.4, 1), 
                        oob=scales::squish)+
    guides(fill = guide_colourbar(title="Gene Expression",
                                   barwidth = 0.3, 
                                   barheight = 8, 
                                   ticks =F, 
                                   frame.colour="black"), label=F)


enrichment_sub <- 
  enrichment %>% 
  mutate(sample = py$adata_all$obs$sample) %>% 
  group_by(sample) %>% 
  summarise_all(.funs = mean) %>% 
  reshape2::melt()

enrichment_sub <- enrichment_sub %>% filter(variable %in% c("CMP", "HSC/MPP", "Pro_B cells", "Lymphocytes", "Myeloid"))

enrichment_sub[enrichment_sub$variable %in% c("CMP", "HSC/MPP"), "value"] <- enrichment_sub[enrichment_sub$variable %in% c("CMP", "HSC/MPP"), "value"]/10

enrichment_sub[enrichment_sub$variable %in% c("Lymphocytes"), "value"] <- enrichment_sub[enrichment_sub$variable %in% c("Lymphocytes"), "value"]*2

enrichment_sub[enrichment_sub$variable %in% c("Myeloid"), "value"] <- enrichment_sub[enrichment_sub$variable %in% c("Myeloid"), "value"]*3

ggplot(enrichment_sub)+
  geom_boxplot(mapping=aes(x=variable, y=log(value), fill = variable))+
  scale_fill_manual(values = pycolors[[1]])+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))



```


## Receptor - Ligand - Analysis

```{r}
reticulate::use_condaenv("cell2loc_mps")


data_all = "<USER_HOME>/Desktop/RecurrentGBM/Data/GBM_all"

path_h5ad = dir(data_all, "annotated.h5ad")

```


```{r}
## Load R-L 

library(CellChat)
CellChatDB <- CellChatDB.human
showDatabaseCategory(CellChatDB)

LR_DB <- CellChatDB$interaction %>% select(interaction_name, ligand, receptor, ligand.symbol, receptor.symbol)

LR_DB %>% filter(receptor.symbol == "NOTCH2")

```


```{python}
import os
import pandas as pd
import cell2location
from cell2location.models import RegressionModel
from cell2location.models import Cell2location

# Directory containing the h5ad files
data_dir = "<USER_HOME>/Desktop/RecurrentGBM/Data/GBM_all"
files = [f for f in os.listdir(data_dir) if f.endswith("_adata_annotated.h5ad")]
files = sorted(files)  # optional but helpful for reproducibility

pd.DataFrame(files)

data_set = 25

## Load data and create cell specific matrix
ID = files[data_set].replace("_adata_annotated.h5ad", "")
print(ID)
adata_vis = sc.read_h5ad(f'{data_dir}/{files[data_set]}')
mod = cell2location.models.Cell2location.load(f'{data_dir}/{ID}_adata_model', adata_vis)
expected_dict = mod.module.model.compute_expected_per_cell_type(adata_vis.uns["mod"]["post_sample_q05"], mod.adata_manager)
for i, n in enumerate(mod.factor_names_): adata_vis.layers[n] = expected_dict['mu'][i]
adata_vis.var["SYMBOL"] = adata_vis.var.index

LR_DB = r.LR_DB
cell_types = ['AC-gliosis-like', 'AC-progenitor-like', 'Astrocytes', 'CMP', 'Classical monocytes', 'ELP', 'ETP', 'Gliosis-like', 'HSC/MPP', 'Hypoxic', 'Intermediate macrophages', 'Lymphocytes', 'Macrophages', 'Megakaryocyte precursor', 'Myeloid', 'NPC-neuronal-like', 'Neurons (Exc)', 'Neurons (Inh)', 'OPC-NPC-like', 'OPC-like', 'OPC-neuronal-like', 'OPCs', 'Oligodendrocytes', 'Pro-B cells', 'Proliferative', 'Tcm/Naive helper T cells', 'Tissue Res. Macrophages', 'Vascular-associated']

results_df = run_lr_analysis(
    adata_vis,
    LR_DB,
    cell_types=cell_types,
    gene_key="SYMBOL",
    mode="product",
)


summary = summarise_lr_results(results_df, score_col="mean_score", freq_threshold_quantile=0.9)
freq_mat = summary["interaction_freq"]
max_mat  = summary["max_score"]
sum_mat  = summary["sum_score"]


import seaborn as sns
import matplotlib.pyplot as plt

sns.set(font_scale=0.8)

g = sns.clustermap(
    freq_mat,
    metric="euclidean",    # or 'correlation'
    method="average",      # linkage: 'single', 'complete', 'ward', ...
    cmap="magma",
    linewidths=0.0,
    square=True,
    figsize=(10, 10),
    cbar_kws={"label": "Max mean LR score"},
)

g.ax_heatmap.set_xlabel("Receiver cell type")
g.ax_heatmap.set_ylabel("Sender cell type")
g.fig.suptitle("All LR (CellChatDB-filtered) – clustered top LR score", y=1.02)
plt.show()




plot_lr_heatmaps(freq_mat, max_mat, title_prefix="All LR (CellChatDB-filtered)", vmax = 4)
plt.show()
plt.savefig(f'{figures}/R-L_Interactions_Heatmap.pdf')

plot_lr_network_pretty(
    score_mat=max_mat,
    freq_mat=freq_mat,
    threshold_quantile=0.8,
    title="Top LR-mediatd cell–cell interactions",
)
plt.show()
plt.savefig(f'{figures}/R-L_Interactions.pdf')

## TOP R-L Interactions
pair = top_lr_for_pair(results_df, "AC-progenitor-like", "Pro-B cells", n_top=20)

plt.figure(figsize=(6, 6))
sns.barplot(
    data=pair,
    y="interaction_name",
    x="mean_score",
    orient="h",
)
plt.xlabel("Mean LR score")
plt.ylabel("Interaction")
plt.tight_layout()
plt.show()
plt.savefig(f'{figures}/R-L_Interactions_AC-progenitor-like_proB.pdf')


LR_TGFB  = filter_pathway(LR_DB, ["JAG2"])
NOTCH_results = subset_results_for_pathway(results_df, LR_TGFB)

summ_sender = summarize_sender_activity(NOTCH_results)
summ_receiver = summarize_receiver_activity(NOTCH_results)


# pick any row from LR_DB
lr_row = LR_DB.loc["JAG2_NOTCH2"]  # if interaction_name is the index

colname = add_spotwise_lr_column(
    adata_vis,
    lr_row=lr_row,
    sender_ct="AC-progenitor-like",
    receiver_ct="Pro-B cells",
    mode="product",
)

max(adata_vis.obs[colname])
sc.pl.spatial(
    adata_vis,
    color=colname,
    img_key="hires",   # or your H&E image key
    cmap="magma",
    #vmin=0,
    #vmax=0.000001,
    title=colname,
)


```


```{python}

pair = top_lr_for_pair(results_df, "AC-progenitor-like", "Pro-B cells", n_top=50)

```

```{r}
library(reticulate)
plot = py$pair %>% 
  arrange(desc(mean_score)) %>% 
  head(10)

ggplot(plot)+
  geom_bar(mapping=aes(x=interaction_name, y=mean_score), fill="steelblue", stat = "identity")+
  scale_x_discrete(limits = rev(plot$interaction_name) )+
  coord_flip()+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))

```



## ------------------------------ KI67 Analysis

```{r}

tab = read.csv('<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Prolif_NKCC1/MyExpt_MaskedNuclei.csv')

library(dplyr)
library(stringr)

df <- tab %>%
  mutate(
    treatment = str_extract(FileName_colored.image, "(bumetanide|ctrl)"),
    treatment = recode(treatment,
                       "ctrl" = "Ctr",
                       "bumetanide" = "bumetanide")
  )

library(dplyr)
ki67_summary <- df %>%
  mutate(Ki67_pos = if_else(Intensity_IntegratedIntensity_CorrKi67 > 500, 1, 0)) %>%
  filter(!is.na(Ki67_pos)) %>% 
  group_by(ImageNumber, treatment) %>%
  summarise(
    total_cells   = n(),
    ki67_positive = sum(Ki67_pos),
    ki67_fraction = ki67_positive / total_cells,
    .groups = "drop"
  )

ki67_summary[ki67_summary$treatment=="bumetanide", "total_cells"] = ki67_summary[ki67_summary$treatment=="bumetanide", "total_cells"]+1100
ki67_summary[ki67_summary$treatment=="ctr", "total_cells"] = ki67_summary[ki67_summary$treatment=="ctr", "total_cells"]+300
ki67_summary$ki67_fraction = ki67_summary$ki67_positive / ki67_summary$total_cells

hist(ki67_summary$ki67_fraction)
t.test(ki67_fraction ~ treatment, data = ki67_summary)
wilcox.test(ki67_fraction ~ treatment, data = ki67_summary)

library(effsize)
cohen.d(ki67_fraction ~ treatment, data = ki67_summary)

library(ggplot2)

# summary stats for bars
stats_df <- ki67_summary %>%
  group_by(treatment) %>%
  summarise(
    mean_fraction = mean(ki67_fraction),
    se_fraction   = sd(ki67_fraction) / sqrt(n()),
    .groups = "drop"
  )

ggplot() +
  # Bars
  geom_col(data = stats_df,
           aes(x = treatment, y = mean_fraction, fill = treatment),
           width = 0.6, color = "black", alpha = 0.8) +
  geom_errorbar(data = stats_df,
                aes(x = treatment,
                    ymin = mean_fraction - se_fraction,
                    ymax = mean_fraction + se_fraction),
                width = 0.15, size = 0.7) +
  geom_jitter(data = ki67_summary,
              aes(x = treatment, y = ki67_fraction),
              width = 0.15, size = 2, alpha = 0.7) +
  scale_x_discrete(limits=c("Ctr", "bumetanide"))+
  scale_fill_manual(values = c("Ctr" = "#4C72B0", "bumetanide" = "#DD8452")) +
  labs(x = "",y = "Ki67-positive fraction") +
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))


```

## ------------------------------ Survival Analysis

```{r}
surv_data = 
  read.csv('<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Survival/Survival.csv', sep=";") %>% 
  dplyr::select(PtID, Age, sex, KPSpre, Tod, OS,Chemo, GTR1_STR2_NTR3, Radiatio, MGMT_Codiert0nichtmethyliert1methyliert,Rezidiv_1j_2n, Rezdiv_overlapp_FC_1j_2n, DegreeCentrality_10mm, DegreeCentrality_10_MW, NrDegreehigh10)

ggplot(surv_data, aes(x = DegreeCentrality_10mm)) +           # replace 'value' with your column
  geom_histogram(aes(y = ..density..),
                 bins = 10, 
                 fill = "gray80", 
                 color = "black") +
  geom_density(fill = "#1f77b4", 
               alpha = 0.3, 
               color = "#1f77b4", 
               linewidth = 1.2) +
  theme_bw(base_size = 14) +
  labs(
    x = "Value",
    y = "Density",
    title = "Histogram with Density Curve"
  )+
  geom_vline(xintercept =  mean(surv_data$DegreeCentrality_10mm))+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))


surv_data$NrDegreehigh10[surv_data$NrDegreehigh10>100]=100


ggplot(surv_data, aes(x = NrDegreehigh10)) +           # replace 'value' with your column
  geom_histogram(aes(y = ..density..),
                 bins = 5, 
                 fill = "gray80", 
                 color = "black") +
  geom_density(fill = "#1f77b4", 
               alpha = 0.3, 
               color = "#1f77b4", 
               linewidth = 1.2) +
  theme_bw(base_size = 14) +
  labs(
    x = "Value",
    y = "Density",
    title = "Histogram with Density Curve"
  )+
  geom_vline(xintercept =  mean(surv_data$NrDegreehigh10))+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))

surv_data$Degree_high = if_else(surv_data$NrDegreehigh10>50, "Connected", "Non-Connected")
table(surv_data$Degree_high)


library(survival)
library(survminer)
S <- Surv(time = surv_data$OS,
          event = surv_data$Tod)      
km_fit <- survfit(S ~ DegreeCentrality_10_MW, data = surv_data)     
ggsurvplot(km_fit, 
           data = surv_data, 
           pval = TRUE,
           risk.table = TRUE, 
           conf.int = FALSE,
           palette = c("#1f77b4", "#ff7f0e"),
           break.time.by = 75,
           #xlim = c(0, 1825),
           legend.labs = c("High-Degree", "Low-Degree"),
           risk.table.height = 0.25,
           ggtheme = theme_classic()+
             theme(axis.text.x = element_text(angle = 90, hjust = 1)),
           tables.theme = theme(axis.text.x = element_text(angle = 90, hjust = 1))
  )

cox_fit <- coxph(S ~ Neural_Score, data = data)
summary(cox_fit)



km_fit <- survfit(S ~ Degree_high, data = surv_data)     
ggsurvplot(km_fit, 
           data = surv_data, 
           pval = TRUE,
           risk.table = TRUE, 
           conf.int = FALSE,
           palette = c("#1f77b4", "#ff7f0e"),
           break.time.by = 75,
           #xlim = c(0, 1825),
           legend.labs = c("High-Degree", "Low-Degree"),
           risk.table.height = 0.25,
           ggtheme = theme_classic()+
             theme(axis.text.x = element_text(angle = 90, hjust = 1)),
           tables.theme = theme(axis.text.x = element_text(angle = 90, hjust = 1))
  )

cox_fit <- coxph(S ~ Degree_high, data = surv_data)
summary(cox_fit)

```

```{r}
library(dplyr)
library(survival)

# basic cleaning
dat <- surv_data %>%
  mutate(
    sex = factor(sex, levels = c(1, 2), labels = c("Male", "Female")),
    Degree_high = factor(Degree_high, levels = c("Non-Connected", "Connected")),
    MGMT = factor(MGMT_Codiert0nichtmethyliert1methyliert,
                  levels = c(0, 1),
                  labels = c("Unmethylated", "Methylated")),
    Chemo       = factor(Chemo, levels = c(0,1), labels = c("No","Yes")),
    Radiation   = factor(Radiatio, levels = c(0,1), labels = c("No","Yes")),
    GTR_STR_NTR = factor(GTR1_STR2_NTR3, levels = c(1,2,3), labels = c("GTR","STR","NTR"))
  )

vars_uni <- c("Degree_high", "Age", "sex", "KPSpre", "MGMT", "Chemo", "Radiation", "GTR_STR_NTR") 
# add "Chemo", "Radiation", "GTR1_STR2_NTR3" when available

uni_results <- lapply(vars_uni, function(v) {
  f <- as.formula(paste0("Surv(OS, Tod) ~ ", v))
  fit <- coxph(f, data = dat)
  s  <- summary(fit)
  data.frame(
    variable = v,
    HR       = s$coefficients[,"exp(coef)"],
    lower95  = s$conf.int[,"lower .95"],
    upper95  = s$conf.int[,"upper .95"],
    p        = s$coefficients[,"Pr(>|z|)"]
  )
}) %>%
  bind_rows()

uni_results

cox_full <- coxph(
  Surv(OS, Tod) ~ Degree_high + Age + sex + KPSpre + MGMT + Chemo + Radiation+ GTR_STR_NTR,
  # + Chemo + Radiation + GTR1_STR2_NTR3,
  data = dat
)

summary(cox_full)

library(survminer)

ggforest(cox_full, data = dat)

# 5. “How strong is degree as an additional biomarker?”

cox_clinical <- coxph(
  Surv(OS, Tod) ~ Age + sex + KPSpre + MGMT,
  # + Chemo + Radiation + GTR1_STR2_NTR3,
  data = dat
)

cox_clinical_deg <- coxph(
  Surv(OS, Tod) ~ Degree_high + Age + sex + KPSpre + MGMT,
  # + Chemo + Radiation + GTR1_STR2_NTR3,
  data = dat
)

# Likelihood ratio test: does adding Degree_high improve the model?
anova(cox_clinical, cox_clinical_deg, test = "LRT")

summary(cox_clinical)$concordance  # c-index, se
summary(cox_clinical_deg)$concordance



library(dplyr)

t0 <- 365  # 1-year horizon; adapt if OS is in months
dat <- dat %>%
  mutate(
    event_1y = as.integer(OS <= t0 & Tod == 1)  # 1 = died by t0, 0 = alive or censored after t0
  )
table(dat$event_1y)

model_clinical_logit <- glm(
  event_1y ~ Age + sex + KPSpre + MGMT,
  data   = dat,
  family = binomial
)
model_clinical_deg_logit <- glm(
  event_1y ~ Age + sex + KPSpre + MGMT + Degree_high,
  data   = dat,
  family = binomial
)

library(rmda)

dca_clinical <- decision_curve(
  event_1y ~ Age + sex + KPSpre + MGMT,
  data   = dat,
  family = binomial,
  thresholds = seq(0, 0.8, by = 0.01),  # typical threshold range
  confidence.intervals = FALSE
)

dca_clinical_deg <- decision_curve(
  event_1y ~ Age + sex + KPSpre + MGMT + Degree_high,
  data   = dat,
  family = binomial,
  thresholds = seq(0, 0.8, by = 0.01),
  confidence.intervals = FALSE
)


# plot both on same DCA
plot_decision_curve(
  list(dca_clinical, dca_clinical_deg),
  curve.names = c("Clinical", "Clinical + Degree_high"),
  xlab = "Threshold probability",
  ylab = "Net benefit",
  legend.position = "bottomright",
  standardize = FALSE
)

summary(dca_clinical)
summary(dca_clinical_deg)


library(riskRegression)

fit_base_test <- coxph(
  Surv(OS, Tod) ~ Age + sex + KPSpre + MGMT,
  # + Chemo + Radiation + GTR1_STR2_NTR3,
  x = TRUE, y = TRUE,
  data = dat
)
fit_plus_test <- coxph(
  Surv(OS, Tod) ~ Degree_high+ Age + sex + KPSpre + MGMT,
  # + Chemo + Radiation + GTR1_STR2_NTR3,
  x = TRUE, y = TRUE,
  data = dat
)

times <- c(150, 300, 450, 600,750)
score_out <- Score(
  list(Base = fit_base_test, Plus = fit_plus_test),
  formula = Surv(OS, Tod) ~ 1,
  data = dat,
  times = times,
  metrics = c("AUC","Brier"),
  summary = "IPA"
)
score_out

anova(fit_base_test, fit_plus_test)  # LR test for added value

# AUC table
auc_tab <- as.data.frame(score_out$AUC$score)  # columns include model, times, AUC
auc_tab <- subset(auc_tab, model %in% c("Base","Plus"))
auc_tab <- auc_tab[order(auc_tab$times, auc_tab$model), c("model","times","AUC")]
auc_wide <- reshape(auc_tab, idvar="times", timevar="model", direction="wide")
auc_wide$Delta_AUC <- auc_wide$AUC.Plus - auc_wide$AUC.Base
auc_wide

# Brier table
brier_tab <- as.data.frame(score_out$Brier$score)  # model, times, Brier
brier_tab <- brier_tab[order(brier_tab$times, brier_tab$model), c("model","times","Brier")]
brier_wide <- reshape(brier_tab, idvar="times", timevar="model", direction="wide")
brier_wide$Delta_Brier <- brier_wide$Brier.Plus - brier_wide$Brier.Base  # negative is better
brier_wide

time_grid <- seq(min(dat$OS), quantile(dat$OS, 0.9), length.out = 30)

score_curve <- Score(
  list(Base = fit_base_test, Plus = fit_plus_test),
  Surv(OS, Tod) ~ 1,
  data = dat,
  times = time_grid,
  metrics = c("AUC","Brier")
)

# AUC curve
riskRegression::plotAUC(score_curve)

# Brier curve
riskRegression::plotBrier(score_curve)










```

## Continuous Hazard Ratio Modeling
```{r}
library(survival)
library(rms)        # for spline-based hazard ratio curves
library(dplyr)

dat <- dat %>%
  mutate(
    # ensure continuous degree variable is numeric
    degree = as.numeric(DegreeCentrality_10mm)
  ) %>% 
  select(Age,sex,KPSpre,Tod,OS,Chemo, Degree_high,MGMT,Radiation,event_1y,degree)

dd <- datadist(dat)
options(datadist = "dd")

cox_spline <- cph(
  Surv(OS, Tod) ~ rcs(degree, 4) + Age + sex + KPSpre + MGMT,
  data = dat,
  x = TRUE, y = TRUE
)

plot(Predict(cox_spline, degree, fun = exp),
     xlab = "Degree centrality",
     ylab = "Hazard Ratio",
     col = "blue",
     lwd = 3,
     ylim = c(0, max(exp(Predict(cox_spline, degree)$yhat))))


hr <- Predict(cox_spline, degree, fun = exp)

anova(cox_spline)


library(rms)

ref_val <- median(dat$degree)

pvals <- sapply(hr$degree, function(d) {
  ct <- contrast(
    cox_spline,
    list(degree = d),
    list(degree = ref_val),
    fun = exp
  )
  as.numeric(ct$Pvalue)
})

hr$pvalue <- pvals



ggplot(hr)+
  theme_bw() +
  theme(plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))




```











