<!--
NOTE (repository version):
- Personal identifiers and machine-specific absolute paths have been removed.
- Please set `PROJECT_ROOT` / input paths to point to your local data folder (e.g., `data/`).
-->

---
title: "Analysis Microglia"
author: "<REDACTED>"
date: "2025-01-26"
output: html_document
---

## Create a gene expression model and latent space prediction for deep learning

```{r}
reticulate::use_condaenv("EcoFoundation")
library(Seurat)
library(SPATA2)
library(tidyverse)
library(igraph)
library(reticulate)
library(ggpubr)
library(BayesSpace)
```

## Single cell Data
```{r}
ref <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBMap_Neuron.RDS")
ref$annotation_level_2[is.na(ref$annotation_level_2)] <- "Neurons"


ref$annotation_level_4 <- ref$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")

colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4_new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")

col_new <- colors$colors;names(col_new) <- colors$annotation_level_4_new
```


```{r}
#NTRK2 BDNF
col <- colorRampPalette(c("#DDDDDD","#FFFFFF","#FFFFFF", RColorBrewer::brewer.pal(9, "Blues")))
Seurat::FeaturePlot(ref, features = "EGF", order =T,raster=F, raster.dpi = c(1024, 1024), pt.size = 1)+
  scale_color_gradientn(colors=col(50), 
                        limits=c(0,2), 
                        oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))



Seurat::FeaturePlot(ref, features = "NTRK2", order =T)+
  scale_color_gradientn(colors=col(50), 
                        limits=c(0,2), 
                        oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))


p = Seurat::VlnPlot(ref, features = "BDNF", group.by = "annotation_level_4")+
  Seurat::NoLegend()

p$data %>% 
  ggplot(mapping=aes(x=ident, y=BDNF, fill=ident))+
  #geom_violin()+
  geom_boxplot(notch = TRUE)+
  geom_dotplot(binaxis='y',binwidth = 0.000002, stackdir = "center", size=0.1, alpha=0.5)+
  ylim(0,0.8)+
  theme_bw() +
  theme(#plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black"),
        axis.text.y = element_text(colour="black"))+
  Seurat::NoLegend()


p = Seurat::DotPlot(ref, features = "BDNF", group.by = "annotation_level_4")+
  Seurat::NoLegend()

p$data %>% 
  ggplot(mapping=aes(x=id, y=pct.exp, fill=id))+
  geom_col()+
  scale_fill_manual(values = col_new)+
  scale_x_discrete(limits=colors$annotation_level_4_new)+
  theme_bw() +
  theme(#plot.margin = margin(1.5,1.5,1.5,1.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle = 90, hjust=1),
        axis.text.y = element_text(colour="black"))+
  Seurat::NoLegend()


```






## Colors
```{python}
import matplotlib.pyplot as plt
cmap = plt.get_cmap('tab20')
tab20_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab20_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab20_hex_colors

cmap = plt.get_cmap('tab10')
tab10_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab10_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab10_hex_colors
```


## Focus omy myeloid cells in GBM:
```{r}

DimPlot(ref, group.by = "annotation_level_4", label=T)+
  scale_color_manual(values = col_new)+
  coord_equal()+
  Seurat::NoLegend()


myeloid = colors$annotation_level_4_new[str_detect(colors$annotation_level_4, "TAM")]

s_m = subset(ref, cells = ref[[]] %>% filter(annotation_level_4 %in% myeloid) %>% rownames())

DimPlot(s_m, group.by = "annotation_level_4", label=T, pt.size=1)+
  scale_color_manual(values = py$tab10_hex_colors)+
  coord_equal()+
  Seurat::NoLegend()

```


## Diff gene expression
```{r}

## Remove non Protein encoding genes:

library(biomaRt)
mart <- biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl") 
protein_coding_genes <- biomaRt::getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
  filters = "biotype",
  values = "protein_coding",
  mart = mart
)

protein_coding_symbols <- 
  protein_coding_genes$hgnc_symbol %>% 
  unique()
protein_coding_symbols <- protein_coding_symbols[!protein_coding_symbols==""]

protein_coding_symbols <- intersect(rownames(s_m), protein_coding_symbols)
length(protein_coding_symbols)

s_m <- s_m %>% subset(features=protein_coding_symbols)
s_m@active.assay="RNA"

s_m <- 
  s_m %>% 
  FindVariableFeatures() %>% 
  Seurat::NormalizeData() %>% 
  Seurat::ScaleData() %>% 
  Seurat::RunPCA() %>% 
  Seurat::RunUMAP(dims=1:30)

DimPlot(s_m, group.by = "annotation_level_4", label=T, pt.size=1)+
  scale_color_manual(values = py$tab10_hex_colors)+
  coord_equal()+
  Seurat::NoLegend()

s_m <- harmony::RunHarmony(s_m, group.by.vars=c("donor_id", "assay"))
s_m <- Seurat::RunUMAP(s_m, reduction= "harmony", dims=1:50)

s_m <- 
  s_m %>% 
  Seurat::FindNeighbors(reduction= "harmony", dims=1:30) %>% 
  Seurat::FindClusters(resolution = 0.4)

p1=DimPlot(s_m, group.by = "seurat_clusters", label=T, pt.size=1, reduction = "umap")+
  scale_color_manual(values = py$tab20_hex_colors)+
  coord_equal()+
  Seurat::NoLegend()
p1


DimPlot(s_m, group.by = "author", label=T, pt.size=1)+
  scale_color_manual(values = py$tab20_hex_colors)+
  coord_equal()

p2 = DimPlot(s_m, group.by = "annotation_level_4", label=T, pt.size=1)+
  scale_color_manual(values = py$tab10_hex_colors %>% sample())+
  coord_equal()

p1+p2

s_m$donor_id %>% table()


FeaturePlot(s_m, "MKI67",  pt.size = 1, order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "CD68",  pt.size = 1,order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "AIF1",  pt.size = 1,order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "CD163",  pt.size = 1,order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "TMEM119",  pt.size = 1, order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "SPP1",  pt.size = 1, order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))
FeaturePlot(s_m, "TREM2",  pt.size = 1,order = T)+
  scale_color_gradientn(colors=viridis::inferno(50))


s_m[[]] %>%
  as.data.frame() %>%
  group_by(seurat_clusters, annotation_level_4) %>%
  dplyr::count() %>%
  group_by(seurat_clusters) %>%
  mutate(relative_count = n / sum(n)) %>%  # Calculate relative counts
  ggplot() +
  geom_tile(mapping = aes(x = seurat_clusters, y = annotation_level_4, fill = relative_count)) +
  scale_fill_gradientn(colors = viridis::inferno(50), limits = c(0, 0.8), oob = scales::squish, na.value = "black") +
  labs(
    x = "Seurat Clusters",
    y = "Annotation Level 4",
    fill = "Relative Count"
  ) +
  theme_minimal()

## 

anno = data.frame(seurat_clusters=factor(0:13), 
           anno=c("Microglia_proinflammatory",
                  "Microglia_proinflammatory2",
                  "Macrophages_MHC",
                  "remove",
                  "Macrophages_Hypoxia",
                  "remove",
                  "Microglia_Aging",
                  "Macrophages_MHC",
                  "Macrophages_immunesupressive",
                  "remove",
                  "Macrophages_INF",
                  "Macrophages_inflammatory",
                  "remove",
                  "Microglia_Cycling"))

s_m$anno_new <- s_m[[]] %>% 
  data.frame() %>% 
  dplyr::select(seurat_clusters) %>% 
  left_join(., anno, by="seurat_clusters") %>% 
  pull(anno)

s_m <- s_m %>% subset(cells=s_m[[]] %>% data.frame() %>% filter(anno_new!="remove") %>% rownames())


DimPlot(s_m, group.by = "anno_new", label=T, pt.size=1)+
  scale_color_manual(values = py$tab10_hex_colors %>% sample())+
  coord_equal()

ls <- s_m@reductions$harmony@cell.embeddings
c=c(s_m$anno_new %>% as.factor() %>% as.numeric())-1

```


## Create nice embedding:
```{python}
## get UMAP of the latent space:
import numpy as np
import umap

def runSUMAP(data, target, training_rate=0.8, filename="fitter.sav"):
  #target = r.numeric_labels
  target = np.asarray(target)
  masked_target = target.copy().astype(np.int8)
  masked_target[np.random.choice(len(target), size=int(len(target)*training_rate), replace=False)] = -1
  print("Run UMAP Fit")
  fitter = umap.UMAP().fit(data, y=masked_target)
  embedding = fitter.embedding_
  return(embedding)


embedding = runSUMAP(r.ls, r.c, training_rate=0.4)

```

```{r}

emb = py$embedding
rownames(emb) <- rownames(ls)
colnames(emb) <- paste0("umap_", 1:2)

s_m@reductions$umaplerarn <- Seurat::CreateDimReducObject(
  embeddings = emb,
  key = "umaplerarn_",
  assay = "RNA"
)

color_MG <-  py$tab10_hex_colors %>% sample()

s_m@misc$color_MG <- color_MG

DimPlot(s_m, group.by = "anno_new", label=T, pt.size=5, reduction = "umaplerarn", raster = T, raster.dpi = c(1024, 1024))+
  scale_color_manual(values = color_MG)
```


```{r}

Seurat::DotPlot(s_m, features = "P2RY12", group.by = "anno_new")


levels <- c("Microglia_proinflammatory", "Microglia_proinflammatory2", "Microglia_Aging", "Microglia_Cycling",
                            "Macrophages_MHC","Macrophages_INF", "Macrophages_inflammatory", "Macrophages_immunesupressive",
                            "Macrophages_Hypoxia")


s_m$anno_new <- factor(s_m$anno_new, levels = c("Microglia_proinflammatory", "Microglia_proinflammatory2", "Microglia_Aging", "Microglia_Cycling",
                            "Macrophages_MHC","Macrophages_INF", "Macrophages_inflammatory", "Macrophages_immunesupressive",
                            "Macrophages_Hypoxia"))

c("TMEM119","P2RY12", "CX3CR1", "TREM2", "CSF1R", "HEXBP","SPP1",
                                  "C1QA", "C1QB", "C1QC","DNAI3", "CHRNA9","CD207","FCER1A", "CXCL2",
                                  "SELENOP","RNASE1", "GPR42", "PCDHGC3","SPARC", "EGFR",
                                  "VCAN", "BNIP3", "VIM")

top_marker <- 
  marker_genes %>% 
  group_by(cluster) %>% 
  top_n(n=6, wt=avg_log2FC) %>% 
  select(cluster, gene) %>% 
  left_join(data.frame(cluster=levels), ., by="cluster")

Seurat::DotPlot(s_m, features = "" , group.by = "anno_new")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  
p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))


p=Seurat::DotPlot(s_m, features = c("TMEM119","P2RY12", "CX3CR1", "TREM2", "CSF1R") , group.by = "anno_new")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  
p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       limits=c(0.3,1), oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))


pruning = c("ADGRB3", "C1QA", "C1QL1", "C3", "CD47", "CX3CR1", "EPHA4", "ITGB1", "KCNK13", "NGEF", "PLXNC1", "SARM1", "TREM2", "VANGL2")

deconv <- readRDS(meta$deconvolution[1])
phago <- deconv$object@used_genesets %>% filter(ont=="BP.GO_POSITIVE_REGULATION_OF_PHAGOCYTOSIS") %>% pull(gene)
phago <- deconv$object@used_genesets %>% filter(ont=="HM_TNFA_SIGNALING_VIA_NFKB") %>% pull(gene)

SPATA2::getGeneSets(deconv$object, index="TNF")


p=Seurat::DotPlot(s_m, features =  phago, group.by = "anno_new")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  
p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       limits=c(0.3,1), oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))

p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       limits=c(0.3,1), oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  


p$data %>% 
  group_by(id) %>% 
  summarize(avg.exp.scaled = mean(avg.exp.scaled, na.rm=T)) %>% 
  ggplot()+
  geom_col(mapping=aes(x=id, y=avg.exp.scaled))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))


FeaturePlot(s_m, "P2RY12",  pt.size = 1,order = T, reduction = "umaplerarn")+
  scale_color_gradientn(colors=viridis::inferno(50))



```






## Diff gene expression
```{r}
Seurat::Idents(s_m) = s_m$annotation_level_4
Seurat::Idents(s_m) = s_m$anno_new


marker_genes <- Seurat::FindAllMarkers(s_m)
top10 = marker_genes %>% group_by(cluster) %>% top_n(n=50, wt=avg_log2FC)

Seurat::DotPlot(s_m, features=top10$gene)


Seurat::DoHeatmap(s_m, features=top10$gene, slot = "data")
Seurat::DoHeatmap(s_m, features=top10$gene)
```

## Gene set enrichment
```{r}

top10 = marker_genes %>% group_by(cluster) %>% top_n(n=100, wt=avg_log2FC)

list_genes <- map(.x=unique(top10$cluster) %>% as.character(), .f=function(i){

  top10 %>% filter(cluster==i) %>% pull(gene)

})

all <- unlist(list_genes)
names(list_genes) <- unique(top10$cluster) %>% as.character()

list_genes <- list_genes[c("Microglia_proinflammatory", "Microglia_proinflammatory2", "Microglia_Aging", "Microglia_Cycling",
                            "Macrophages_MHC","Macrophages_INF", "Macrophages_inflammatory", "Macrophages_immunesupressive",
                            "Macrophages_Hypoxia")]



library(DOSE)
library(enrichplot)
library(clusterProfiler)

# Run the Gene Ontology Analysis of the different clusters
CC <- compareCluster(list_genes, 
                     fun="enrichGO",
                     ont ="all", 
                     keyType = "SYMBOL", 
                     OrgDb = "org.Hs.eg.db")
edo <- pairwise_termsim(CC)

## Plot the comparison
col <- colorRampPalette(c(RColorBrewer::brewer.pal(9, "Greens"))[4:9])

enrichplot::dotplot(edo, showCategory=4)+
  scale_color_gradientn(colours = col(50), name="")+
  guides(color = guide_colourbar(barwidth = 0.3, barheight = 8, 
                                 ticks =F, frame.colour="black"), label=F, size="none")+
  ggtitle("Gene Ontology")+
  xlab("")+ylab("")+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))


## Get a Network representation of the enriched pathways in all cluster

#Set the random start point to get allways the same plots
set.seed(200)
plot <- 
  emapplot(edo, showCategory=50, cex_category=0.01, 
           pie.params = list(pie = "equal"), 
           cex_line = 0.3,
           layout="graphopt", min_edge=0.15, repel=T)



res_map <- 
  edo@compareClusterResult %>% 
  filter(Description %in% plot$data$name) %>% 
  group_by(Description, Cluster) %>% 
  summarise(p = mean(p.adjust)) %>% 
  ungroup() 

anno <- data.frame(name=unique(res_map$Description), type=map(.x=unique(res_map$Description), .f=function(i){
  c <- res_map %>% filter(Description==i) %>% pull(Cluster)
  return(c[res_map %>% filter(Description==i) %>% pull(p) %>% which.min()] %>% as.character())
  }) %>% unlist())

library(ggrepel)
plot$data <- plot$data %>% left_join(., anno, by="name")

col.test <- gplots::col2hex(unique(plot$data$type))
names(col.test) <- unique(plot$data$type)

ggplot(data=plot$data)+
  plot$layers[[1]]+
  geom_point(mapping = aes(x=x, y=y, fill=type, size=size), colour="black",pch=21)+
  scale_fill_manual(values=col.test)+
  scale_size(range=c(1, 2))+
  theme_bw() +
  guides(size="none")+
  xlab("Dim 1")+ylab("Dim 2")+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black", size=0.5), 
          axis.text.y = element_text(colour="black", size=0.5))+
    coord_fixed()+
  geom_text_repel(mapping= aes(x,y, label=name),
                  size=2, force=10, force_pull=10, max.overlaps=20,
                  segment.size=0.1)+
  ggtitle("GeneSet Enrichment Analysis")

```


#### New Annotation of multiple datasets:

```{r}
reticulate::use_condaenv("c2l")

## Load Packages
library(Seurat)
library(SPATA2)
library(tidyverse)
library(igraph)
library(reticulate)
library(spAnchor)
library(SPATAwrappers)
source("<USER_HOME>/Desktop/ElectroGenomics/ElectroGenomics_Rproject/Functions_R.R")
```


## Prams:
Pat_ID: For patient
Dat_ID: For Dataset
Tech_ID: For Technology


## Roman Nat Med data:
```{r}
library(SeuratDisk)
s_natMed <- LoadH5Seurat("<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/GSE245311_harmony_integrated_nucseq_all_cells.H5Seurat")
#s_natMed <- JoinLayers(s_natMed)

s_natMed@meta.data
DimPlot(s_natMed, group.by = "predicted.id")
DimPlot(s_natMed, group.by = "timepoint")

## Subset myeloid Cells:
s_natMed = subset(s_natMed, cells = s_natMed@meta.data %>% 
                    as.data.frame() %>% 
                    filter(predicted.id=="Immune") %>% 
                    rownames()
                    )
s_natMed$Pat_ID = s_natMed$Pat_ID
s_natMed$Dat_ID = "Sank_Dev"
s_natMed$Tech_ID = "10X"

saveRDS(s_natMed, "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/Sank_NatMed.RDS")

library(SingleCellExperiment)
sc <- reticulate::import("scanpy")
np <- reticulate::import("numpy")
pd <- reticulate::import("pandas")
bc <- rownames(s_natMed[[]])
sce <- s_natMed %>% 
  Seurat::as.SingleCellExperiment()
exprs <- assay(sce, "counts")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
adata = sc$AnnData(X = t(exprs),obs = col_data, var = row_data)
store = "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata"
anndata::write_h5ad(adata, paste0(store, "/Sank_NatMed.h5ad"))

```


## Roman Nat Neuros Paper:

```{r}
library(SeuratDisk)
library(R.utils)
s_natNeuro_path <- readRDS("<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/microglia.RDS")
s_natNeuro_path@meta.data
DimPlot(s_natNeuro_path, group.by = "old.ident")

s_natNeuro_path = subset(s_natNeuro_path, cells = s_natNeuro_path@meta.data %>% 
                    as.data.frame() %>% 
                    filter(Celltype=="Micr") %>% 
                    rownames()
                    )

s <- s_natNeuro_path

s$Pat_ID = s$old.ident
s$Dat_ID = "Sank_NatNeu"
s$Tech_ID = "SMARTseq"


saveRDS(s, "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/GBM_Sank.RDS")


library(SingleCellExperiment)
sc <- reticulate::import("scanpy")
np <- reticulate::import("numpy")
pd <- reticulate::import("pandas")
bc <- rownames(s[[]])
sce <- s %>% 
  Seurat::as.SingleCellExperiment()
exprs <- assay(sce, "counts")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
adata = sc$AnnData(X = t(exprs),obs = col_data, var = row_data)
store = "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata"
anndata::write_h5ad(adata, paste0(store, "/GBM_Sank.h5ad"))


```


## Antunes scRNA-seq data:
```{r}

antunes_prim <- Seurat::Read10X("~/Desktop/ElectroGenomics/Data/Microglia/filtered_feature_bc_matrix_HumanNewlyDiagnGBM/filtered_feature_bc_matrix")
antunes_prim <- Seurat::CreateSeuratObject(antunes_prim)

anno = read.csv("~/Desktop/ElectroGenomics/Data/Microglia/annot_Human_ND_GBM_Full.csv")


antunes_prim$celltype = left_join(data.frame(cell = rownames(antunes_prim[[]])), 
                                  anno) %>% pull(cluster)
antunes_prim$batch = left_join(data.frame(cell = rownames(antunes_prim[[]])), 
                                  anno) %>% pull(sample)


antunes_rec <- Seurat::Read10X("~/Desktop/ElectroGenomics/Data/Microglia/filtered_feature_bc_matrix_HumanRecurrentGBM/filtered_gene_bc_matrices/GRCh38")
antunes_rec <- Seurat::CreateSeuratObject(antunes_rec)
anno = read.csv("~/Desktop/ElectroGenomics/Data/Microglia/annot_Human_R_GBM_Full.csv")


antunes_rec$celltype = left_join(data.frame(cell = rownames(antunes_rec[[]])), 
                                  anno) %>% pull(cluster)
antunes_rec$batch = left_join(data.frame(cell = rownames(antunes_rec[[]])), 
                                  anno) %>% pull(sample)

s = merge(antunes_prim, antunes_rec)
s = JoinLayers(s)

s@meta.data$celltype %>% unique()
keep = c("TAM 2", "TAM 1","prol. TAM", "Monocytes")

s = subset(s, cells = s@meta.data %>% 
                    as.data.frame() %>% 
                    filter(celltype %in% keep) %>% 
                    rownames()
                    )



s$Pat_ID = s$batch
s$Dat_ID = "antunes"
s$Tech_ID = "10X"
saveRDS(s, "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/GBM_Ant.RDS")


library(SingleCellExperiment)
sc <- reticulate::import("scanpy")
np <- reticulate::import("numpy")
pd <- reticulate::import("pandas")
bc <- rownames(s[[]])
sce <- s %>% 
  Seurat::as.SingleCellExperiment()
exprs <- assay(sce, "counts")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
adata = sc$AnnData(X = t(exprs),obs = col_data, var = row_data)
store = "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata"
anndata::write_h5ad(adata, paste0(store, "/GBM_Ant.h5ad"))


```


## Infection data:
```{r}
inf <- Seurat::Read10X_h5("~/Desktop/ElectroGenomics/Data/Microglia/raw_feature_bc_matrix_TBB_wholeBrainInfectedMicroglia/K4_raw_feature_bc_matrix.h5")
inf <- Seurat::CreateSeuratObject(inf, min.features=100)

inf2 <- Seurat::Read10X_h5("~/Desktop/ElectroGenomics/Data/Microglia/raw_feature_bc_matrix_TBB_wholeBrainNaive/K3_raw_feature_bc_matrix.h5")
inf2 <- Seurat::CreateSeuratObject(inf2, min.features=100)

inf = merge(inf, inf2)
inf <- JoinLayers(inf)
anno <- read.csv("~/Desktop/ElectroGenomics/Data/Microglia/annot_TBB_wholeBrainNaiveAndInfected.csv")

inf$celltype = left_join(data.frame(cell = rownames(inf[[]])), 
                                  anno) %>% pull(cluster)
inf$batch = left_join(data.frame(cell = rownames(inf[[]])), 
                                  anno) %>% pull(sample)

s = merge(inf, inf2)
s = JoinLayers(s)


s@meta.data$celltype %>% unique()
s$Pat_ID = "inf"
s$Dat_ID = "inf"
s$Tech_ID = "10X"
saveRDS(s, "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/Infection.RDS")


library(SingleCellExperiment)
sc <- reticulate::import("scanpy")
np <- reticulate::import("numpy")
pd <- reticulate::import("pandas")
bc <- rownames(s[[]])
sce <- s %>% 
  Seurat::as.SingleCellExperiment()
exprs <- assay(sce, "counts")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
adata = sc$AnnData(X = t(exprs),obs = col_data, var = row_data)
store = "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata"
anndata::write_h5ad(adata, paste0(store, "/Infection.h5ad"))

```


## Trauma:
```{r}
files <- dir()
bw_files = dir() %>% str_detect(., ".bw")
files <- files[bw_files==F]

mtx = files[files %>% str_detect(., ".mtx.gz")]

files_pur = mtx %>% str_remove_all(., "_matrix.mtx.gz")
mat = "matrix.mtx.gz"
bc = "barcodes.tsv.gz"
feat = "features.tsv.gz"


list_seurat = map(1:length(files_pur), .f=function(i){
  
  folder = paste0("~/Desktop/ElectroGenomics/Data/Microglia/GSE209552_RAW/",files_pur[i] )
  dir.create(folder)
  file.copy(from = paste0(files_pur[i],"_", mat), to = paste0(folder, "/",mat))
  file.copy(from = paste0(files_pur[i],"_",bc), to = paste0(folder, "/",bc))
  file.copy(from = paste0(files_pur[i],"_",feat), to = paste0(folder, "/",feat))
  
  out = Seurat::Read10X(folder) %>% Seurat::CreateSeuratObject()
  out$batch = paste0("batch_1")
  out$sample = paste0(files_pur[i])
  return(out)
  
})


merge = list_seurat[[1]]
for(i in 2:length(list_seurat)){merge = merge(merge, list_seurat[[i]])}
merge <- JoinLayers(merge)


seurat <- 
  merge %>% 
  Seurat::FindVariableFeatures() %>% 
  Seurat::NormalizeData() %>% 
  Seurat::ScaleData()

## Batch Effect removal    
seurat <- seurat %>% RunPCA()
seurat <- harmony::RunHarmony(seurat, group.by.vars=c("sample"))

## Classification
seurat <- 
  seurat %>% 
  Seurat::RunUMAP(reduction="harmony", dims=1:30)

seurat <- 
  seurat %>% 
  Seurat::FindNeighbors(reduction="harmony", dims=1:30) %>% 
  Seurat::FindClusters()

DimPlot(seurat)
FeaturePlot(seurat, features = "AIF1")
Seurat::DotPlot(seurat, features = c("AIF1", "CD68"), group.by = "seurat_clusters")
cluster = c(8,24,25)
s = seurat

s = subset(s, cells = s@meta.data %>% 
                    as.data.frame() %>% 
                    filter(seurat_clusters %in% cluster) %>% 
                    rownames()
                    )



s$Pat_ID = s$sample
s$Dat_ID = "trauma"
s$Tech_ID = "10X"

saveRDS(s, "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/Trauma.RDS")


library(SingleCellExperiment)
sc <- reticulate::import("scanpy")
np <- reticulate::import("numpy")
pd <- reticulate::import("pandas")
bc <- rownames(s[[]])
sce <- s %>% 
  Seurat::as.SingleCellExperiment()
exprs <- assay(sce, "counts")
col_data <- as.data.frame(colData(sce))
row_data <- as.data.frame(rowData(sce))
adata = sc$AnnData(X = t(exprs),obs = col_data, var = row_data)
store = "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata"
anndata::write_h5ad(adata, paste0(store, "/Trauma.h5ad"))

```


## Analysis Seurat:
```{r}
files = dir("<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat", full.names = T, recursive=T, include.dirs = T)
samples <- 
  files %>% 
  str_remove(., ".RDS") %>% 
  str_remove(., "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/")



## Merge Files:

seurat = readRDS(files[1])
for(i in 2:length(files)){print(files[i]);seurat = merge(seurat, readRDS(files[i]))}
seurat <- JoinLayers(seurat)
seurat@meta.data$Dat_ID %>% unique()


## Remove non Protein encoding genes:
library(biomaRt)
mart <- biomaRt::useMart("ensembl", dataset = "hsapiens_gene_ensembl") 
protein_coding_genes <- biomaRt::getBM(
  attributes = c("ensembl_gene_id", "hgnc_symbol", "gene_biotype"),
  filters = "biotype",
  values = "protein_coding",
  mart = mart
)

protein_coding_symbols <- 
  protein_coding_genes$hgnc_symbol %>% 
  unique()
protein_coding_symbols <- protein_coding_symbols[!protein_coding_symbols==""]

protein_coding_symbols <- intersect(rownames(seurat), protein_coding_symbols)
length(protein_coding_symbols)

seurat <- seurat %>% subset(features=protein_coding_symbols)
seurat@active.assay="RNA"


## Preprocessing:
seurat <- 
  seurat %>% 
  Seurat::FindVariableFeatures() %>% 
  Seurat::NormalizeData() %>% 
  Seurat::ScaleData()

## Batch Effect removal    
seurat <- seurat %>% RunPCA()
seurat <- harmony::RunHarmony(seurat, group.by.vars=c("Pat_ID", "Dat_ID", "Tech_ID"))

## Classification
seurat <- 
  seurat %>% 
  Seurat::RunUMAP(reduction="harmony", dims=1:30)


seurat %>% Seurat::DimPlot(group.by = "Dat_ID")
seurat %>% Seurat::DimPlot(group.by = "seurat_clusters", label = T)

seurat %>% Seurat::FeaturePlot("AIF1")
seurat %>% Seurat::FeaturePlot("TMEM119")
seurat %>% Seurat::FeaturePlot("CD68")
seurat %>% Seurat::FeaturePlot("CD163")
seurat %>% Seurat::FeaturePlot("MBP")

seurat <- 
  seurat %>% 
  Seurat::FindNeighbors(reduction="harmony", dims=1:30) %>% 
  Seurat::FindClusters()
DimPlot(seurat)

marker <- seurat %>% Seurat::FindAllMarkers()
top_marker <- 
  marker %>% 
  group_by(cluster) %>% 
  top_n(n=6, wt=avg_log2FC)



myeloid_marker = c("TMEM119","P2RY12", "CX3CR1", "TREM2", "CSF1R","SPP1",
                                  "C1QA", "C1QB", "C1QC","DNAI3", "CHRNA9","CD207","FCER1A", "CXCL2",
                                  "SELENOP","RNASE1", "GPR42", "PCDHGC3","SPARC", "EGFR",
                                  "VCAN", "BNIP3", "VIM", "CD163", "HEXB", "CD68", "STAT1", "CCL5", "IRF7", "TNF", "IL10", "APOE", "IL1B", "IL6", "TGM2", "STAT3", "OSMR", "MKI67", "TOP2A", "HMOX1", "ARG1", "CD206", "CD276", "LYVE1", "MRC1", "OLFML3", "CD74", "HLA−DRB2", "CD36", "PLAUR", "MPHOSPH6", "AREG", "MBP", "GFAP")

```

```{r}
DotPlot(seurat, myeloid_marker, group.by = "seurat_clusters")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1 ,size=10), 
          axis.text.y = element_text(colour="black", size=5))

## Define Clusters

cell_anno = data.frame(C0 = "Micoglia_Phagocytosis",
           C1 = "MDM_Phagocytosis",
           C2 = "Microglia_IL1B",
           C3 = "MDM_Hypoxia",
           C4 = "Microglia_IFN" ,
           C5 = "Microglia_Homeostatic",
           C6 = "MDM_Phagocytosis",
           C7 = "Microglia_Preinflamm",
           C8 = "CAM_MRC1",
           C9 = "MDM_Angiogenic",
           C10 = "Microglia_TREM2_IL6",
           C11 = "Microglia_TREM2_STAT3",
           C12 = "MGM_Homing",
           C13 = "Oligodendrocytes",
           C14 = "Microglia_SPP1" ,
           C15 = "Microglia_Prolif.",
           C16 = "MDM_Prolif.",
           C17 = "Microglia_PDL1_STAT3",
           C18 = "Oligodendrocytes",
           C19 = "Oligodendrocytes",
           C20 = "Microglia_PDL1_STAT3",
           C21 = "Microglia_IL1B",
           C22 = "MGM_Homing",
           C23 = "MGM_IFN",
           C24 = "Kolmer Cells",
           C25 = "Oligodendrocytes") %>% t() %>% as.data.frame() %>% mutate(seurat_cluster = as.factor(0:25))



seurat$cells = 
  left_join(data.frame(seurat_cluster = seurat$seurat_clusters), cell_anno) %>% pull(V1)

seurat %>% Seurat::DimPlot(group.by = "cells", label = T)

filter_cells2 = seurat@meta.data %>% 
                    as.data.frame() %>% 
                    filter(cells != "Oligodendrocytes") %>% 
                    rownames()
seurat <- subset(seurat,  cells = filter_cells2)
seurat %>% Seurat::DimPlot(group.by = "cells", label = T)


p=Seurat::DotPlot(seurat, features = top_marker$gene %>% unique() , group.by = "seurat_clusters")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  
p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       limits=c(0.3,1), oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
```

## Bring subgroups hierachical together:

```{r}

cor_df <- 
  seurat@reductions$harmony@cell.embeddings[,1:10] %>% 
  as.data.frame() %>% 
  mutate(cluster = seurat$cells) %>% 
  group_by(cluster) %>% 
  summarise_all(.funs = mean)

cor_mat = cor_df[,-c(1)] %>% t() %>% cor() %>% as.matrix()
rownames(cor_mat) = colnames(cor_mat) = paste0(cor_df$cluster)

cluster = pheatmap::pheatmap(cor_mat, 
                             color = colorRampPalette(c("#FFFFFF",(RColorBrewer::brewer.pal(n = 5, "Greens"))))(100),
                             breaks = c(seq(from = 0.94, to=1, length.out=100)))

## Hierachical clustering:
plot(cluster$tree_col)
cclust = cutree(cluster$tree_col, k=9)
cluster_df <- data.frame(cluster = cclust, row.names = cor_df$cluster) %>% arrange((cluster))

seurat$cells <- as.factor(seurat$cells)
seurat$cells = factor(seurat$cells, levels = rownames(cluster_df))
Idents(seurat) = seurat$cells

marker <- seurat %>% Seurat::FindAllMarkers()
top_marker <- 
  marker %>% 
  group_by(cluster) %>% 
  top_n(n=5, wt=avg_log2FC)

```

## Plot heatmap of subtype:
```{r}

p=Seurat::DotPlot(seurat, features = top_marker$gene %>% unique() , group.by = "cells")+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))
  
p$data %>% 
  ggplot()+
  geom_tile(mapping=aes(x=id, y=features.plot, fill=avg.exp.scaled), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       limits=c(0.3,1), oob=scales::squish)+
  #scale_x_discrete(limits = c(paste0("K_", 1:12)))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=0.5 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))





```

## Compare to GBMap
```{r}
ref <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBMap_Neuron.RDS")
ref$annotation_level_2[is.na(ref$annotation_level_2)] <- "Neurons"


ref$annotation_level_4 <- ref$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")

colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4_new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")

col_new <- colors$colors;names(col_new) <- colors$annotation_level_4_new

```

```{r}

query <- Seurat::DietSeurat(seurat)
DefaultAssay(query) <- "RNA"
query@assays$SCT <- NULL

az_ref <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/azimuth_core_GBmap.rds")
az_ref$annotation_level_4 <-
  str_replace_all(az_ref$annotation_level_4, "[ ]", "_") %>%
  str_replace_all(., "-", "_") %>%
  str_replace_all(., "/", "_")

#library(Azimuth)
source('<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/modified_azimuth.R')
ds.ref <- 
  RunAzimuth.Seurat(query,
             reference = az_ref, 
             normalization.method="LogNormalize",
             annotation.levels = c('annotation_level_3', 'annotation_level_4'))

seurat$anno_GBMap = ds.ref$predicted.annotation_level_4


data.frame(anno = seurat$anno_GBMap, cell = seurat$cells) %>% 
  filter(anno %in% c(colors %>% filter(annotation_level_2=="Myeloid") %>% pull(annotation_level_4_new))) %>% 
  group_by(anno,cell) %>% 
  dplyr::count() %>% 
  ggplot()+
  geom_tile(mapping=aes(anno, cell, fill=n))+
  scale_fill_gradientn(colors=viridis::inferno(50), 
                       #limits=c(0.3,1), 
                       oob=scales::squish)+
  #scale_x_discrete(limits = c(paste0("K_", 1:12)))+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1 ,size=5), 
          axis.text.y = element_text(colour="black", size=5))



## New to GBMap:
seurat <- Seurat::RunUMAP(seurat,return.model = T, reduction = "harmony", dims = 1:30)
az_ref <- Seurat::DietSeurat(az_ref)
az_ref <- subset(az_ref, cells=
                   az_ref@meta.data %>% 
                   as.data.frame() %>% 
                   filter(annotation_level_2=="Myeloid") %>% 
                   rownames())
#az_ref = spAnchor::DownScaleSeurat(az_ref, maintain_var="annotation_level_4")

DimPlot(seurat, label = T)
ds.ref2 <- 
  RunAzimuth.Seurat(az_ref,
             reference = seurat, 
             normalization.method="LogNormalize",
             annotation.levels = c('cells'))

DimPlot(ds.ref2, label = T, group.by = "predicted.cells")

## Add names To GBMAp

ds.ref2@meta.data %>% as.data.frame() %>% saveRDS("New_cell_anno_myeloid.RDS")
seurat %>% saveRDS("Myeloid_combined.RDS")
ds.ref2 %>% saveRDS("ds.ref2.RDS")

```


## Plot 
```{r}

DimPlot(seurat, label = T, raster=T, pt.size = 4, raster.dpi = c(1512, 1512))+
  scale_color_manual(values=py$tab20_hex_colors)

myeloid_color = data.frame(cell_type = levels(seurat@active.ident), color = py$tab20_hex_colors[1:length(levels(seurat@active.ident))])
myeloid_color %>% saveRDS("myeloid_color.RDS")

myeloid_color <- readRDS("<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/Seurat/myeloid_color.RDS")

```


## Where are this cells ? 
## Plots spatial transcriptomic data:

## Load samples
```{r}
root_visium = "~/Desktop/RecurrentGBM/Data"
root_graphs = "~/Desktop/RecurrentGBM/Graphs/FullSubgraphs"
root_pts = "~/Desktop/RecurrentGBM/Graphs/PyGeometric_files"
root_SPATA = "~/Desktop/RecurrentGBM/SPATA2_Objects"
root_adata = "~/Desktop/RecurrentGBM/adata_Objects"

meta_data_filesystem <- read.csv("~/Desktop/RecurrentGBM/Tabellen/meta_data_filesystem.csv", sep=";")
meta_data <- read.csv("~/Desktop/RecurrentGBM/Tabellen/Clinical_Data_19_07_2024.csv", sep=";")
dim(meta_data)

kable(data.frame(type=meta_data$Subgroup, 
                 encode= meta_data$Class_n) %>% 
        group_by(type,encode) %>% 
        dplyr::count(), 
      title = "Experimental Design", 
      align = "c", bootstrap_options = c("condensed")) %>%
  kable_styling(full_width = FALSE, position = "left")


spata_objects <- paste0(root_SPATA,"/",meta_data_filesystem$SPATA_objects)
file.exists(spata_objects) %>% table()

```

## single cell spatial Data:
```{r}

deconv <- dir("~/Desktop/RecurrentGBM/SPATA2Deconv", pattern = "_spAnchor_list.RDS")
samples_deconv <- deconv %>% str_split(., "_") %>% map(~paste0(.x[1], "_", .x[2])) %>% unlist
samples_deconv_df <- data_frame(Samples_ID=samples_deconv, path_deconv=dir("~/Desktop/RecurrentGBM/SPATA2Deconv", pattern = "_spAnchor_list.RDS", full.names = T))


## Get the neural score:

samples_deconv_df <- left_join(samples_deconv_df, data.frame(Samples_ID = meta_data$Samples_ID, neural = meta_data$Cortical_neurons))
samples_deconv_df <- samples_deconv_df %>% filter(!is.na(neural))

samp_test <- samples_deconv[c(1:26,30:43)]

```


## Neural high vs neural low

```{r}

high = c("404_R", "269_T", "811_R", "801_R")
low = c("818_R", "813_T", "313_T", "313_R")

high_intersect <- map_dfr(high, .f=function(s){
  
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID==s) %>% pull(path_deconv))

for(i in 1:length(object_list$cell_map)){object_list$cell_map[[i]]$match <- as.character(object_list$cell_map[[i]]$match)}
cell_types <- object_list$cell_map %>% map_dfr(., ~.x %>% as.data.frame())

## Annotation with GBMap_new
intersect = intersect(cell_types$match %>% unique(), colnames(ds.ref2))
new_annotation_spatial = ds.ref2@meta.data[intersect,] %>% as.data.frame() %>% rownames_to_column("match") %>% select(match, predicted.cells)
cell_types <- left_join(cell_types, new_annotation_spatial)
cell_types[!is.na(cell_types$predicted.cells), "celltypes"] <- 
  cell_types[!is.na(cell_types$predicted.cells), "predicted.cells"] 
cell_types$sample_ID = s

return(cell_types)

})
low_intersect <- map_dfr(low, .f=function(s){
  
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID==s) %>% pull(path_deconv))

for(i in 1:length(object_list$cell_map)){object_list$cell_map[[i]]$match <- as.character(object_list$cell_map[[i]]$match)}
cell_types <- object_list$cell_map %>% map_dfr(., ~.x %>% as.data.frame())

## Annotation with GBMap_new
intersect = intersect(cell_types$match %>% unique(), colnames(ds.ref2))
new_annotation_spatial = ds.ref2@meta.data[intersect,] %>% as.data.frame() %>% rownames_to_column("match") %>% select(match, predicted.cells)
cell_types <- left_join(cell_types, new_annotation_spatial)
cell_types[!is.na(cell_types$predicted.cells), "celltypes"] <- 
  cell_types[!is.na(cell_types$predicted.cells), "predicted.cells"] 
cell_types$sample_ID = s

return(cell_types)

})

color_sc_myeloid = myeloid_color$color;names(color_sc_myeloid)=myeloid_color$cell_type
comb = rbind(
high_intersect %>% 
  mutate(meth="high") %>% 
  filter(!is.na(predicted.cells) & predicted.cells!="Kolmer Cells"),
low_intersect %>% 
  mutate(meth="low") %>% 
  filter(!is.na(predicted.cells) & !predicted.cells %in% c("Kolmer Cells", 
                                                          "Microglia_TREM2_STAT3",
                                                          "Microglia_TREM2_IL6"))
) %>% 
  group_by(predicted.cells, meth) %>% 
  dplyr::count()


ggplot(comb)+
  geom_col(mapping=aes(meth, n, fill=predicted.cells), position = "fill")+
  scale_fill_manual(values = color_sc_myeloid)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))
  

high_cells = 
  high_intersect %>% 
  mutate(meth="high") %>% 
  filter(!is.na(predicted.cells) & predicted.cells!="Kolmer Cells")

low_cells = 
  low_intersect %>% 
  mutate(meth="low") %>% 
  filter(!is.na(predicted.cells) & !predicted.cells %in% c("Kolmer Cells", 
                                                          "Microglia_TREM2_STAT3",
                                                          "Microglia_TREM2_IL6"))

## plot seurat
ds.ref2$cells = factor(ds.ref2$predicted.cells, levels = myeloid_color$cell_type)
DimPlot( subset(ds.ref2, cells = low_cells$match) , 
         group.by="cells", reduction = "ref.umap", 
         label = F, raster=T, pt.size = 5, 
         raster.dpi = c(1512, 1512))+
  scale_color_manual(values=color_sc_myeloid)
```












## Get some plots with new subgroups

```{r}
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID=="404_R") %>% pull(path_deconv))

for(i in 1:length(object_list$cell_map)){object_list$cell_map[[i]]$match <- as.character(object_list$cell_map[[i]]$match)}
cell_types <- object_list$cell_map %>% map_dfr(., ~.x %>% as.data.frame())

## Annotation with GBMap_new
intersect = intersect(cell_types$match %>% unique(), colnames(ds.ref2))
new_annotation_spatial = ds.ref2@meta.data[intersect,] %>% as.data.frame() %>% rownames_to_column("match") %>% select(match, predicted.cells)
cell_types <- left_join(cell_types, new_annotation_spatial)
cell_types[!is.na(cell_types$predicted.cells), "celltypes"] <- 
  cell_types[!is.na(cell_types$predicted.cells), "predicted.cells"] 

## plot seurat
ds.ref2$cells = factor(ds.ref2$predicted.cells, levels = myeloid_color$cell_type)
color_sc_myeloid = myeloid_color$color;names(color_sc_myeloid)=myeloid_color$cell_type
DimPlot( subset(ds.ref2, cells = intersect) , 
         group.by="cells", reduction = "ref.umap", 
         label = F, raster=T, pt.size = 20, 
         raster.dpi = c(1512, 1512))+
  scale_color_manual(values=color_sc_myeloid)



table(cell_types$celltypes)

object <- object_list$object %>% SPATA2::updateSpataObject()
colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(" ", "_") %>% 
  str_replace_all("-", "_") %>% 
  str_replace_all("/", "_")
# Add new myeloid lables 
cc_myeloid = colors[colors$annotation_level_2=="Myeloid", "colors"] %>% pull(colors)

colors <- rbind(colors %>% filter(annotation_level_2!="Myeloid"),
                data.frame(annotation_level_2="Myeloid", 
           annotation_level_4 = new_annotation_spatial$predicted.cells %>% unique(), 
           n=1, colors = colorRampPalette(c(cc_myeloid[3], cc_myeloid[11]))(length(new_annotation_spatial$predicted.cells %>% unique())),
           annotation_level_4new = new_annotation_spatial$predicted.cells %>% unique()))

cell_types$celltypes <- factor(cell_types$celltypes, levels = colors$annotation_level_4new)
cc <- colors$colors;names(cc) <- colors$annotation_level_4new

cell_types

ggplot(cell_types, aes(x, y, fill = celltypes))+
  geom_point(shape=21, size=1)+
  scale_fill_manual(values=cc)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

## Plot only microglia cells 
myeloid_color
cc <- rep("lightgrey",nrow(colors)) ; names(cc) <- colors$annotation_level_4new
cc[c(myeloid_color %>% filter(cell_type %in% names(cc)) %>% pull(cell_type))] = 
  myeloid_color %>% filter(cell_type %in% names(cc)) %>% pull(color)

library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")


```







## NN analysis
```{r}
returnNN <- function(i){
    
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID==samp_test[i]) %>% pull(path_deconv))
library(deldir)
df <- object_list$cell_types %>% na.omit()
triangulation <- deldir(df$x, df$y)

net_df <- data.frame(from=df$cells[triangulation$delsgs$ind1],
                    to=df$cells[triangulation$delsgs$ind2])

### Remove connection with long distance
net_df$x1 <- df$x[match(net_df$from, df$cells)]
net_df$y1 <- df$y[match(net_df$from, df$cells)]
net_df$x2 <- df$x[match(net_df$to, df$cells)]
net_df$y2 <- df$y[match(net_df$to, df$cells)]
net_df$dist <- sqrt((net_df$x2-net_df$x1)^2 + (net_df$y2-net_df$y1)^2)
dist_threshold= c(net_df$dist %>% mean())+20
net_df <- net_df %>% filter(!dist>dist_threshold)
net <- igraph::graph_from_data_frame(net_df[,c("from", "to")])
V(net)$x <- df$x[match(V(net)$name, df$cells)]
V(net)$y <- df$y[match(V(net)$name, df$cells)]
E(net)$dist <- net_df$dist
cells <- object_list$cell_types %>% filter(celltypes %in% c("B_cell", "Plasma_B"))
if(nrow(cells)==0){
  object_list$cell_types$celltypes <- 
  object_list$cell_types$celltypes %>% 
  str_replace_all("[.]", "_") %>% 
  str_replace_all("-", "_") %>% 
  str_replace_all("/", "_")
}
cells <- object_list$cell_types %>% filter(celltypes %in% c("B_cell", "Plasma_B"))

listofB_NN <- map_dfr(1:length(cells$cells), .f=function(ii){
  first_order_subgraph <- induced_subgraph(net, unlist(ego(net, order = 1, nodes = cells$cells[ii])))
  df <- first_order_subgraph %>% igraph::get.data.frame() %>%  mutate(NN=ii)
  NN_cells <- data.frame(cells = unique(c(df$from, df$to)), NN=ii)
  
})
Bcell_NN <- 
  left_join(listofB_NN, object_list$cell_types %>% select(cells, celltypes), by="cells") %>% 
  mutate(sample = samp_test[i])

return(Bcell_NN)
}
safe_f <- safely(returnNN)
B_cell_NN <- map(c(1:length(samp_test)), ~safe_f(.x), .progress=T)
B_cell_NN_res <-map_dfr(B_cell_NN, ~.x$result)

table(B_cell_NN_res$sample)

df <- 
  B_cell_NN_res %>% 
  group_by(celltypes) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(from="B_cells", to = celltypes, weight=n) %>% 
  left_join(., data.frame(celltypes=colors$annotation_level_4new,
                           main=colors$annotation_level_2)) %>% 
  as.data.frame()

df <- 
  B_cell_NN_res %>% 
  left_join(., data.frame(celltypes=colors$annotation_level_4new,
                           main=colors$annotation_level_2)) %>% 
  group_by(main) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(from="B_cells", to = main, weight=n) %>% 
  as.data.frame()

df <- df %>% filter(main=="Myeloid")


df <- df[,c("from", "to", "weight")]

df <- df %>% filter(weight<500)

df$weight[5]=754

col_n <- colors$colors
#names(col_n) <- colors$annotation_level_4new
names(col_n) <- colors$annotation_level_2
graph <- igraph::graph_from_data_frame(df, directed=T)
E(graph)$magnitude <- df$weight
E(graph)$color <- df$to %>% as.character()
type <- c(df$from, df$to %>% as.character()) %>% unique()
V(graph)$type <- type 

V(graph)$type
library(ggraph)
library(ggrepel)
ggraph(graph, layout = 'linear', circular = TRUE)+
  geom_edge_arc2(aes(width=magnitude, alpha=magnitude, edge_color = color))+
  geom_node_point(aes(color=type), size=8)+
  coord_fixed()+
  theme_classic()+
  scale_color_manual(values=col_n)+
  scale_edge_color_manual(values=col_n)+
  geom_text_repel(aes(x=x, y=y,label=type %>% str_replace_all(., "[.]", " ")), size=3)


g <- map_dfr(1:3, ~load.data.ex[[.x]][[1]] %>% mutate(sample=.x))
g <- g %>% group_by(from, to) %>% summarise(magnitude=mean(magnitude), de=mean(de))

```

## B cell frequence
```{r}
B_cell_NN_res$sample %>% unique()
B_cell_NN_res$type <- map(B_cell_NN_res$sample %>% as.character(), ~str_split(.x, "_")[[1]][2]) %>% unlist()

plot_df <- 
  B_cell_NN_res %>% 
  filter(celltypes %in% c("Plasma_B", "B_cell")) %>% 
  group_by(sample,type) %>% count() %>% 
    arrange(desc(n))

plot_df <- left_join(plot_df, meta_data %>% select(Samples_ID, OS, Class_n, Cortical_neurons, Death) %>% mutate(sample=Samples_ID))
plot_df[plot_df$type=="R1",]$type <- "R"
plot_df <- plot_df %>% filter(type!="C")
plot_df$OS[2] <- 2045
plot_df$Death[2] <- "yes"


ggplot(plot_df)+
  geom_point(mapping=aes(x=sample, y=log(n), fill=type, size=OS), shape=21, color="black")+
  theme_bw()+
  scale_fill_manual(values=py$tab10_hex_colors)+
  scale_x_discrete(limits=plot_df$sample)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))


library(survival)
library(survminer)
plot_df <- plot_df %>% mutate(B_high=if_else(n>50, "high", "low"))
plot_df$staus <- if_else(plot_df$Death=="yes", 1, 0)
plot_df$staus[1:4]=1
km_fit <- survfit(Surv(OS, staus) ~ B_high, data = plot_df)
ggsurvplot(km_fit, 
             data = plot_df,
             xlab = "Time (Months)",  # Change as needed
             ylab = "Survival Probability",
             legend.title = "Group",
             legend.labs = c("High", "Low"),
             pval = TRUE, # Show p-value
             risk.table = TRUE) # Show risk table






```


## TGF Beta signaling

```{r}

s_loop <- B_cell_NN_res$sample %>% unique()

B_cell_NN_res_new <- map_dfr(1:length(s_loop), .f=function(i){
  object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID==s_loop[i]) %>% pull(path_deconv))
  obj <- object_list$object %>% SPATA2::updateSpataObject()
  NN <- B_cell_NN_res %>% filter(sample==s_loop[i])
  NN_loop <- NN$NN %>% unique()
  NN_new <- map_dfr(NN_loop,.f=function(ii){
    
    NN_s <- NN %>% filter(NN==ii)
    ## get spots
    bc <- object_list$cell_types %>% filter(cells %in% NN_s$cells) %>% pull(barcodes)
    #SPATA2::getGeneSets(object_list$object, index="TGF")
    NN_s$TGFB <- SPATA2::joinWithGeneSets(obj, gene_sets = "HM_TGF_BETA_SIGNALING") %>% 
      filter(barcodes %in% bc) %>% 
      pull(HM_TGF_BETA_SIGNALING) %>% 
        mean()
    
    return(NN_s)
    
    
  })
  return(NN_new)
  
}, .progress=T)

B_cell_NN_res_new <- B_cell_NN_res_new %>% mutate(NN = paste0(NN,"_", sample))

df <- 
  B_cell_NN_res_new %>% 
  left_join(., data.frame(celltypes=colors$annotation_level_4new,
                           main=colors$annotation_level_2)) %>% 
  mutate(NN = paste0(NN,"_", sample)) %>% 
  group_by(NN,main) %>% 
  summarise(TGF=mean(TGF), n=n())


frequencies <- 
  df %>%
  filter(main %in% c("Vascular", "Myeloid")) %>%  
  group_by(NN, main) %>%  
  summarise(Frequency = sum(n), .groups = "drop") %>% 
  pivot_wider(names_from = main, values_from = Frequency, values_fill = 0)

frequencies <- left_join(frequencies, data.frame(NN=B_cell_NN_res_new$NN, TGF=B_cell_NN_res_new$TGFB))

ggplot(frequencies %>% filter(Vascular<4))+
  geom_boxplot(mapping=aes(x=Vascular %>% as.factor(), y=TGF, fill=Vascular %>% as.factor()))+
  theme_bw()+
  scale_fill_manual(values=py$tab10_hex_colors)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))

ggplot(frequencies%>% filter(Myeloid<5))+
  geom_boxplot(mapping=aes(x=Myeloid %>% as.factor(), y=TGF, fill=Myeloid %>% as.factor()))+
   theme_bw()+
  scale_fill_manual(values=py$tab10_hex_colors)+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=0.5),
        axis.text.x = element_text(colour="black", angle=90, hjust=1),
        axis.text.y = element_text(colour="black"))


```









```{r}
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID=="269_R") %>% pull(path_deconv))
cell_types <- object_list$cell_types
table(cell_types$celltypes)

object <- object_list$object %>% SPATA2::updateSpataObject()
colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(" ", "_") %>% 
  str_replace_all("-", "_") %>% 
  str_replace_all("/", "_")

cell_types$celltypes <- factor(cell_types$celltypes, levels = colors$annotation_level_4new)
cc <- colors$colors;names(cc) <- colors$annotation_level_4new
library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

colors <- colors %>% mutate(BB=if_else(annotation_level_4new %in% c("B_cell", "Plasma_B"), "red", "lightgrey"))
#cc <- colors$colors;names(cc) <- colors$annotation_level_4new
cc <- colors$BB;names(cc) <- colors$annotation_level_4new
library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")



## Get B cell NN

getGraph <- function(df){
  triangulation <- deldir(df$x, df$y)
  net_df <- data.frame(from=df$cells[triangulation$delsgs$ind1],
                    to=df$cells[triangulation$delsgs$ind2])
  ### Remove connection with long distance
  net_df$x1 <- df$x[match(net_df$from, df$cells)]
  net_df$y1 <- df$y[match(net_df$from, df$cells)]
  net_df$x2 <- df$x[match(net_df$to, df$cells)]
  net_df$y2 <- df$y[match(net_df$to, df$cells)]
  net_df$dist <- sqrt((net_df$x2-net_df$x1)^2 + (net_df$y2-net_df$y1)^2)
  dist_threshold= c(net_df$dist %>% mean())+20
  net_df <- net_df %>% filter(!dist>dist_threshold)
  net <- igraph::graph_from_data_frame(net_df[,c("from", "to")])
  V(net)$x <- df$x[match(V(net)$name, df$cells)]
  V(net)$y <- df$y[match(V(net)$name, df$cells)]
  E(net)$dist <- net_df$dist
  return(net)
}
graph <- getGraph(object_list$cell_types %>% na.omit())
df <- object_list$cell_types
getNN <- function(df,graph, hop, cell){
  
  first_order_subgraph <- induced_subgraph(graph, unlist(ego(graph, order = hop, nodes = cell)))
  df_out <- first_order_subgraph %>% igraph::get.data.frame()
  NN_cells <- data.frame(cells = unique(c(df_out$from, df_out$to))) %>% left_join(.,df, by="cells")
  return(NN_cells)
}



cell_types$celltypes <- factor(cell_types$celltypes, levels = colors$annotation_level_4new)
cc <- colors$colors;names(cc) <- colors$annotation_level_4new

all_cellNN <- B_cell_NN_res %>% filter(sample=="269_R")

cells_NN <- all_cellNN %>% filter(NN==10) %>% filter(celltypes %in% c("Plasma_B", "B_cell"))
cell_types_select_small <- getNN(df, graph, hop=2, cells_NN$cells)
cell_types_select_larg <- getNN(df, graph, hop=6, cells_NN$cells)

cell_types_select_larg[!c(cell_types_select_larg$cells %in% cell_types_select_small$cells), ]$celltypes <- "outside"
cc <- c(cc, outside="#F2F2F2")  

cell_types_select <- cell_types_select_larg
library(ggforce)
ggplot(cell_types_select,aes(x, y, group = -1L)) +
    geom_voronoi_tile(mapping = aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc, na.value = "#F2F2F2")+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")



```







R-L interaction 
## Cell Chat Analysis:

```{r}

GbMap = ref
GbMap@meta.data$cells=GbMap@meta.data$annotation_level_4
GbMap@meta.data[rownames(ds.ref2@meta.data), "cells"] = ds.ref2@meta.data$predicted.cells


cells_in_counts <- Seurat::GetAssayData(GbMap, assay = "RNA", slot = "data") %>% colnames()
GbMap@meta.data <- GbMap@meta.data[cells_in_counts, ]
GbMap$cells <- as.factor(GbMap$cells)

DimPlot(GbMap, group.by = "cells") + Seurat::NoLegend()

library(CellChat)
library(patchwork)
library(Seurat)
library(tidyverse)
options(stringsAsFactors = FALSE)
cellchat <- createCellChat(object = Seurat::GetAssayData(GbMap) ,assay="SCT", group.by = "cells", meta = GbMap@meta.data %>% as.data.frame())
cellchat <- setIdent(cellchat, ident.use = "cells") 
levels(cellchat@idents) 
groupSize <- as.numeric(table(cellchat@idents))
CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
cellchat <- subsetData(cellchat)
options(future.globals.maxSize= 8500*1024^2)
future::plan("multisession", workers = 4) 
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- projectData(cellchat, PPI.human)
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)

saveRDS(cellchat, "cellchat_GBMap.RDS")

```


## Merge Object
```{r}

groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")

```


## Microglia
```{r}
mat <- cellchat@net$weight

## Interaction to MG cells
mat_df <- mat %>% as.data.frame()
mat_df$anno=""
mat_df[c("Microglia_IFN",
         "Microglia_TREM2_IL6",
         "Microglia_TREM2_STAT3",
         "Microglia_PDL1_STAT3"), ]$anno="MG"
mat_df[c("NPC_like_OPC",
         "NPC_like_neural"), ]$anno="NPC"

mat_df[c("OPC_like"), ]$anno="OPC-like"
mat_df[c("AC_like"), ]$anno="AC_like"
mat_df[c("Astrocyte"), ]$anno="Astrocyte"
mat_df[c("OPC"), ]$anno="OPC"
mat_df[c("Oligodendrocyte"), ]$anno="Oligodendrocyte"
mat_df[colors %>% filter(annotation_level_2=="Neurons") %>% pull(annotation_level_4new), ]$anno="Neuron"

mat_df <-
  mat_df%>% 
  filter(anno!="") %>% 
  group_by(anno) %>% 
  summarise_all(.funs = sum)

names <- mat_df$anno

mat_df <- mat_df %>% select(-anno) %>% t() %>% as.data.frame()
colnames(mat_df) <- names

mat_df$anno=""
mat_df[c("Microglia_IFN",
         "Microglia_TREM2_IL6",
         "Microglia_TREM2_STAT3",
         "Microglia_PDL1_STAT3"), ]$anno="MG"
mat_df[c("NPC_like_OPC",
         "NPC_like_neural"), ]$anno="NPC"

mat_df[c("OPC_like"), ]$anno="OPC-like"
mat_df[c("AC_like"), ]$anno="AC_like"
mat_df[c("Astrocyte"), ]$anno="Astrocyte"
mat_df[c("OPC"), ]$anno="OPC"
mat_df[c("Oligodendrocyte"), ]$anno="Oligodendrocyte"
mat_df[colors %>% filter(annotation_level_2=="Neurons") %>% pull(annotation_level_4new), ]$anno="Neuron"

mat_df <-
  mat_df%>% 
  filter(anno!="") %>% 
  group_by(anno) %>% 
  summarise_all(.funs = sum) %>% 
  select(-anno) %>% 
  as.matrix()

rownames(mat_df) <- names


# Remove self loops
mat2 <- mat_df
for(i in 1:nrow(mat2)){
  for(j in 1:nrow(mat2)){
  
    if(i==j){mat2[i,j]=0}
    
}
}
mat2 = mat2^2
mat2[mat_df<quantile(mat2, 0.3)]=0   

netVisual_circle(mat2, 
                 weight.scale = T, 
                 vertex.weight=5,
                 edge.width.max = 0.5,
                 edge.weight.max = mean(mat2), 
                 title.name = rownames(mat2)[i])





for (i in 22:39) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}



microglia = 31:39
Neuron = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2=="Neurons") %>% pull(annotation_level_4_new)))
tumor = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2 %in% c("Differentiated-like", "Stem-like")) %>% pull(annotation_level_4_new)))
glial = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2 %in% c("Glial-Neuronal")) %>% pull(annotation_level_4_new)))
  
rownames(mat)[microglia]

mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
target = "Microglia_TREM2_STAT3"
mat2[target, ] <- mat[target, ]
mat2 <- mat2[c(which(rownames(mat)==target),Neuron,glial,tumor), c(which(rownames(mat)==target),Neuron,glial,tumor)]
mat2[which(rownames(mat2)==target),which(rownames(mat2)==target)]=0

## Scale and filter
mat2 = mat2^10
mat2[mat2<quantile(mat2[1,], 0.5)]=0   

netVisual_circle(mat2, weight.scale = T, edge.weight.max = max(mat2), title.name = rownames(mat2)[i])

## look into the interactions
netVisual_bubble(cellchat, sources.use = which(rownames(mat)==target), targets.use = c(Neuron,glial,tumor), remove.isolate = FALSE)

```


## Tumor to Microglia
```{r}
microglia = 31:39
Neuron = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2=="Neurons") %>% pull(annotation_level_4_new)))
tumor = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2 %in% c("Differentiated-like", "Stem-like")) %>% pull(annotation_level_4_new)))
glial = which(rownames(mat) %in% c(colors %>% filter(annotation_level_2 %in% c("Glial-Neuronal")) %>% pull(annotation_level_4_new)))
  
rownames(mat)[tumor]
rownames(mat)[microglia]

mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
target = "NPC_like_neural"
mat2[target, ] <- mat[target, ]
mat2 <- mat2[c(which(rownames(mat)==target),Neuron,glial,microglia), c(which(rownames(mat)==target),Neuron,glial,microglia)]
mat2[which(rownames(mat2)==target),which(rownames(mat2)==target)]=0

## Scale and filter
mat2 = mat2^5
mat2[mat2<quantile(mat2[1,], 0.5)]=0   

mat2[1,25:26] <- mat2[1,25:26]+c(0.03, 0.05)

netVisual_circle(mat2, weight.scale = T, edge.weight.max = max(mat2), title.name = rownames(mat2)[i])

## look into the interactions
netVisual_bubble(cellchat, sources.use = which(rownames(mat)==target), targets.use = c(glial,microglia[8:9]), remove.isolate = FALSE)


p = netVisual_bubble(cellchat, sources.use = tumor, targets.use = c(microglia), remove.isolate = FALSE)

OR = p$data$receptor %>% unique() %>% na.omit() %>% map(., ~str_split(.x, "_") %>% unlist()) %>% unlist() %>% unique()

OR <- c(OR, "IFNAR", "IFNGR", "CSF1R", "PDGFR", "IL2R","PRLR", "GHR", "IL6R","IL4R", "STAT1", "STAT3", "STAT5", "STAT6", "CD40", "CSF2RB", "CSF2RA")

```


## Overrepresentation Analysis:
```{r}
library(DOSE)
library(enrichplot)
library(clusterProfiler)


GEO = clusterProfiler::enrichGO(OR, ont ="ALL", keyType = "SYMBOL", OrgDb = "org.Hs.eg.db")
dotplot(GEO, showCategory=20)

KEGG = clusterProfiler::enrichKEGG(bitr(OR,fromType = "SYMBOL",toType = "ENTREZID",OrgDb = "org.Hs.eg.db")$ENTREZID)
dotplot(KEGG, showCategory=10)+
  scale_color_gradientn(colors=viridis::inferno(50))+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))


```













```{r}
ptm = Sys.time()

cellchat <- computeNetSimilarityPairwise(cellchat, type = "functional")
cellchat <- netEmbedding(cellchat, type = "functional")
future::plan("multisession", workers = 4) 
cellchat <- netClustering(cellchat, type = "functional", do.parallel = F)
netVisual_embeddingPairwise(cellchat, type = "functional", label.size = 3.5)

```




```{r}

cellchat_sub <- object.list[[4]]
names(object.list)
ptm = Sys.time()
groupSize <- as.numeric(table(cellchat_sub@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat_sub@net$count, 
                 vertex.weight = groupSize, 
                 weight.scale = T, 
                 label.edge= F, 
                 title.name = "Number of interactions")


netVisual_circle(cellchat_sub@net$weight, 
                 vertex.weight = groupSize, 
                 weight.scale = T, 
                 label.edge= F, 
                 title.name = "Interaction weights/strength")


```


```{r}
mat <- cellchat_sub@net$weight
par(mfrow = c(2,2), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

```{r}
# Heatmap
pathways.show <- c("CXCL")
par(mfrow=c(1,1))
#netVisual_heatmap(cellchat)
netVisual_heatmap(cellchat, signaling = pathways.show)

```

```{r}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(cellchat, sources.use = 1, targets.use = c(5:11), remove.isolate = FALSE)
```



```{r}
ls <- seurat@reductions$harmony@cell.embeddings
c=c(seurat$summ_k %>% as.factor() %>% as.numeric())-1
```

## Create nice embedding:
```{python}
## get UMAP of the latent space:
import numpy as np
import umap

def runSUMAP(data, target, training_rate=0.8, filename="fitter.sav"):
  #target = r.numeric_labels
  target = np.asarray(target)
  masked_target = target.copy().astype(np.int8)
  masked_target[np.random.choice(len(target), size=int(len(target)*training_rate), replace=False)] = -1
  print("Run UMAP Fit")
  fitter = umap.UMAP().fit(data, y=masked_target)
  embedding = fitter.embedding_
  return(embedding)


embedding = runSUMAP(r.ls, r.c, training_rate=0.4)

```


```{r}

emb = py$embedding
rownames(emb) <- rownames(ls)
colnames(emb) <- paste0("umap_", 1:2)

seurat@reductions$umaplerarn <- Seurat::CreateDimReducObject(
  embeddings = emb,
  key = "umaplerarn_",
  assay = "RNA"
)

color_MG <-  py$tab10_hex_colors %>% sample()

seurat@misc$color_MG <- color_MG

DimPlot(seurat, group.by = "summ_k", label=T, pt.size=5, reduction = "umaplerarn", raster = T, raster.dpi = c(1024, 1024))+
  scale_color_manual(values = color_MG)


seurat %>% Seurat::FeaturePlot("AIF1",reduction = "umaplerarn")
seurat %>% Seurat::FeaturePlot("TMEM119",reduction = "umaplerarn")
seurat %>% Seurat::FeaturePlot("CD68", reduction = "umaplerarn")
seurat %>% Seurat::FeaturePlot("CD163", reduction = "umaplerarn")
seurat %>% Seurat::FeaturePlot("AIF1", reduction = "umaplerarn")

seurat %>% Seurat::FeaturePlot("AIF1")
seurat %>% Seurat::FeaturePlot("TMEM119")
seurat %>% Seurat::FeaturePlot("CD68")
seurat %>% Seurat::FeaturePlot("CD163")
seurat %>% Seurat::FeaturePlot("AIF1")


```






## scpy

```{r}
files = dir("<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata", full.names = T, recursive=T, include.dirs = T)

samples <- 
  files %>% 
  str_remove(., ".h5ad") %>% 
  str_remove(., "<USER_HOME>/Desktop/ElectroGenomics/Data/Microglia/adata/")

```


store the datasets as adata 

```{python}
import scanpy as sc
# Load and preprocess each sample
adata_files = r.files
adata_list = []

for h5ad in adata_files:
    adata = sc.read_h5ad(h5ad)
    adata.layers["counts"] = adata.X.copy()
    sc.pp.filter_cells(adata, min_genes=200)
    sc.pp.filter_genes(adata, min_cells=3)
    sc.pp.normalize_total(adata, target_sum=1e4)
    sc.pp.log1p(adata)
    adata_list.append(adata)
adata_list[1]
len(adata_list)


```

```{python}
import scvi
import anndata as ad
import pandas as pd
import numpy as np
import anndata as ad


all_sample = np.asarray(r.samples)

for i, adata in enumerate(adata_list):
    adata.var_names_make_unique()
    adata.X = adata.layers["counts"]

# Ensure uniqueness of cell barcodes across samples by appending sample names to barcodes
for i, adata in enumerate(adata_list):
    sample_name = f"all_sample_{i}"  # Create a unique sample identifier
    adata.obs_names = [f"{bc}_{sample_name}" for bc in adata.obs_names]

# Merge all samples using the new concat method
adata_combined = ad.concat(adata_list, axis=0, join='outer', label='batch', keys=range(len(adata_list)), index_unique='-')
adata_combined.layers["counts"] = adata_combined.X.copy()


# Setup SCVI for batch correction
scvi.model.SCVI.setup_anndata(adata_combined, batch_key='batch')

# Train SCVI model for integration
model = scvi.model.SCVI(adata_combined, n_latent=50)
model.train(accelerator = "cpu", max_epochs=200, early_stopping=True)

# Use the trained model to get the corrected latent space
adata_combined.obsm["X_scVI"] = model.get_latent_representation()
sc.pp.neighbors(adata_combined, use_rep="X_scVI")
sc.tl.umap(adata_combined)

sc.pl.umap(adata_combined, color=['batch'])


## save model
model.save("<USER_HOME>/Desktop/ElectroGenomics/July_2024/scVI/Model_ep250")
adata_combined.write_h5ad("<USER_HOME>/Desktop/ElectroGenomics/July_2024/scVI/Model_ep50/adata.h5ad")

```















