<!--
NOTE (repository version):
- Personal identifiers and machine-specific absolute paths have been removed.
- Please set `PROJECT_ROOT` / input paths to point to your local data folder (e.g., `data/`).
-->

---
title: "Analysis MEA data"
author: "<REDACTED>"
date: "2025-02-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Input packages:
```{r}
reticulate::use_condaenv("cell2loc_mps")
library(readxl)
library(patchwork)
library(tidyverse)
library(SPATA2)
library(igraph)
root_out = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/Analysis_DHH"
root_SPATA = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/data/proc"
root_MEA = "<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/UKA_MEA/"
meta <- readRDS("<USER_HOME>/Desktop/ElectroGenomics/Aachen EG/Electrogenomics_newdata/Analysis_DHH/Meta.RDS")

meta[,2] <- str_replace_all(meta[,2], "dieterhenrikheiland", "<REDACTED>")
meta[,3] <- str_replace_all(meta[,3], "dieterhenrikheiland", "<REDACTED>")
meta[,5] <- str_replace_all(meta[,5], "dieterhenrikheiland", "<REDACTED>")
meta[,6] <- str_replace_all(meta[,6], "dieterhenrikheiland", "<REDACTED>")
meta[,7] <- str_replace_all(meta[,7], "dieterhenrikheiland", "<REDACTED>")
meta[,8] <- str_replace_all(meta[,8], "dieterhenrikheiland", "<REDACTED>")
meta$spata2_new <- paste0(root_out, "/", meta$samples, "_spata2.RDS")



colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4_new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")



```


## Colors
```{python}
import matplotlib.pyplot as plt
cmap = plt.get_cmap('tab20')
tab20_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab20_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab20_hex_colors

cmap = plt.get_cmap('tab10')
tab10_colors = cmap.colors

# Converting RGB colors to hexadecimal format
def rgb_to_hex(rgb):
    return "#{:02x}{:02x}{:02x}".format(int(rgb[0]*255), int(rgb[1]*255), int(rgb[2]*255))

tab10_hex_colors = [rgb_to_hex(color) for color in tab20_colors]
tab10_hex_colors
```

Samples die keinen Tumor haben !!

S24_06_CT_01_spata2
S24_06_CT_02_spata2


## Create SPATA files
```{r}
i=1
for(i in 1:nrow(meta)){
  
  raw <- readRDS(meta$raw_data[i])
  scale=1
  object <- SPATA2::initiateSpataObject_CountMtr(coords_df = raw$coords,
                                                 count_mtr=raw$counts,
                                                 image = raw$image$mea_overlay_orient,
                                                 sample_name = meta$samples[i]
                                                 )
  object <- object %>% flipImage(axis="h")
  plotSurface(object)
  
  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))
  
}
meta$spata2_new <- paste0(root_out, "/", meta$samples, "_spata2.RDS")


## Add cytospace:
for(i in 1:nrow(meta)){
  object <- readRDS(meta$spata2_new[i])
  object@spatial[[1]]$cytospace = readRDS(meta$cytospace[i])
  object@spatial[[1]]$graph = readRDS(meta$graph[i])
  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))

}

## Add c2l:
for(i in 1:nrow(meta)){
  object <- readRDS(meta$spata2_new[i])
  output <- paste0(root_out, "/", meta$samples[i] ,"_deconv.csv")
  cell_types <- read.csv(output)
  names(cell_types)[1] <- "barcodes"
  cell_types_df <- cell_types
  names(cell_types_df) <- str_remove(names(cell_types_df), "q05cell_abundance_w_sf_")
  names(cell_types_df) <- names(cell_types_df) %>% str_replace_all(., "[.]", "_")
  object <- object %>% SPATA2::addFeatures(cell_types_df, overwrite = T)

  saveRDS(object, paste0(root_out, "/", meta$samples[i], "_spata2.RDS"))

}

```


## Plot Single cell and mea data:
```{r}
#Colors:
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
i=5
object <- readRDS(meta$spata2_new[i])
plotSurface(object)

object@used_genesets = rbind(data.frame(ont = "connectivity", gene=wnk_spak_genes), object@used_genesets)
limits=c(0.1, 0.8)
p = plotSurface(object, color_by = "wnk_spak_genes", alpha_by="wnk_spak_genes", display_image=F, limits=limits, oob=scales::squish)+
  scale_alpha(limits=limits,oob=scales::squish)
p

cell_types <- object@spatial[[1]]$cytospace


p = plotSurface(object, pt_alpha=0)

library(ggforce)
p=ggplot()+
  geom_voronoi_tile(data =cell_types, mapping=aes(x=x_sc,y=y_sc, group = -1L, fill = CellType), 
                      max.radius = 3, colour = 'black', linewidth=0.1, alpha=1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")
  Seurat::NoLegend()


## Add MEA Grid
library(ggnewscale)
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges

cond <- nodes$drug %>% unique()
drug_select = c(nodes$drug %>% unique())[1]
weight_lim = 0.4

n_inp = nodes[!is.na(nodes$drug), ] %>% filter(.$drug == drug_select)
e_inp = edges %>% filter(.$drug==drug_select) %>% filter(weight>weight_lim)


p+
  ggnewscale::new_scale_color()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=1)+
  geom_segment(data = e_inp %>% arrange(weight), 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight), alpha=0.1)+ 
  scale_size(range=c(0.1,0.3))+
  scale_color_viridis_c(, 
                        limits=c(0, 400), 
                        oob=scales::squish)

ggplot()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=2, color="black")+
  geom_segment(data = e_inp, 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight))+ 
  scale_size(range=c(0.1,1))+
  scale_color_viridis_c(option="B", limits=c(0.3, 0.7), oob=scales::squish)

```

## KNN analysis
```{r}

## Get KNN cells of the MEA spots
library(FNN)
drug_select = c(nodes$drug %>% unique())[1]
n_inp = 
  nodes[!is.na(nodes$drug), ] %>% 
  filter(.$drug == drug_select)

#node_rank = n_inp %>% arrange(desc(degree)) %>% mutate(degree_rank=1:nrow(.)) %>% select(id, degree_rank)
node_rank = n_inp %>% arrange(desc(nonrandom_spikes)) %>% mutate(degree_rank=1:nrow(.)) %>% select(id, degree_rank)


query_points = data.frame(x=n_inp$x, y=n_inp$y)
#data_points = data.frame(x=cell_types$x_sc, y=cell_types$y_sc)
coords = SPATA2::getCoordsDf(object)
data_points = data.frame(x=coords$x, y=coords$y)
k <- 5

# Find k-nearest neighbors of query_points in data_points
knn_result <- get.knnx(data = data_points, query = query_points, k = k)

colors$annotation_level_2[10:13] <- "TAMs"
colors$annotation_level_2[14:21] <- "MGs"
colors$annotation_level_2[55:61] <- "Ex_Neuron"
colors$annotation_level_2[62:66] <- "Inh_Neuron"
colors$annotation_level_2[39:42] <- colors$annotation_level_4_new[39:42]

# Print results
index = knn_result$nn.index
dist = knn_result$nn.dist   

NN = map_dfr(1:nrow(index), .f=function(i){
  print(i)
  data.frame(
    barcodes = coords$barcodes[i],
    degree = n_inp$degree[index[i,]] %>% mean(na.rm=T),
    random = n_inp$nonrandom_spikes[index[i,]]%>% mean(na.rm=T),
    firing_rate = n_inp$firing_rate[index[i,]]%>% mean(na.rm=T)
  )
})

NN <- left_join(NN, joinWithGeneSets(object, gene_sets = "wnk_spak_genes"), by="barcodes")



NN = map_dfr(1:nrow(index), .f=function(i){
  data.frame(
    UniqueCID = cell_types$UniqueCID[index[i,]],
    dist = dist[i,],
    id = n_inp$id[i],
    degree_rank = node_rank %>% filter(id==n_inp$id[i]) %>% pull(degree_rank),
    degree = n_inp$degree[i],
    random = n_inp$nonrandom_spikes[i],
    firing_rate = n_inp$firing_rate[i]
  )
})
NN <- left_join(NN, cell_types, by="UniqueCID")
NN <- left_join(NN, data.frame(CellType = colors$annotation_level_4_new, anno2 = colors$annotation_level_2))


ggplot(NN,mapping = aes(x=wnk_spak_genes, y=degree))+geom_point()+
  geom_smooth(se=F, method="lm")+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))


plot_df <- NN %>% group_by(degree_rank, anno2) %>% count()
plot_df$anno2 <- factor(plot_df$anno2, levels = unique(colors$annotation_level_2))


plot_df
ggplot()+
  geom_col(plot_df, mapping=aes(x=degree_rank, y=n, fill=anno2), position = "fill")+
  #scale_fill_manual(values=cc)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))



plot_df <- plot_df %>% 
                filter(anno2 %in% c("Ex_Neuron", "Inh_Neuron"))
                #filter(degree_rank>100)
ggplot()+
  geom_smooth(plot_df, 
              mapping=aes(x=degree_rank, y=n, color=anno2))+
  #scale_fill_manual(values=cc)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))


plot_df <- plot_df
ggplot()+
  geom_smooth(plot_df, 
              mapping=aes(x=degree_rank, y=n, color=anno2), se=F)+
  #scale_fill_manual(values=cc)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))

```


## Correlation of cell abundance and network parameter

```{r}
library(FNN)
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
getCellCor <- function(object, nodes, cell_types, k=2, drug_select, prams){

  cells <- 
    object %>% 
    joinWithFeatures(features=cell_types)
  
  # subset drug
  n_inp = 
    nodes[!is.na(nodes$drug), ] %>% 
    filter(.$drug == drug_select) %>% 
    group_by(id,x,y) %>% 
    select(id,x,y, {prams}) %>% 
    summarise_all(.funs = mean)
  
  ggplot()+
    geom_point(data = data.frame(x=cells$x, y=cells$y), mapping=aes(x,y))+
    geom_point(data = n_inp[,c("x", "y")],mapping=aes(x,y), color="red")
  
  ## KNN
  knn_result <- get.knnx(data = data.frame(x=cells$x, y=cells$y), query = n_inp[,c("x", "y")] , k = k)
  
  ## Summarize
  NN = map_dfr(1:nrow(knn_result$nn.index), .f=function(i){
      
    cell_mat = 
      cells[knn_result$nn.index[i,], cell_types] %>% colSums() %>% 
      as.data.frame()
    names(cell_mat)[1] <- "fraction"
    cell_mat <- cell_mat %>% rownames_to_column("Cell_type")
    
    cell_mat <- 
      cell_mat %>% 
      mutate(
        dist = mean(knn_result$nn.dist[i,]),
        id = n_inp$id[i])
      
    return(cell_mat)
})
  NN <- NN %>% left_join(., n_inp, by="id")
  
  
  ## run cor:
  all = map_dfc(cell_types, .f=function(ct){
    
    inp=na.omit(NN)
    cor_mat <- inp %>% 
      filter(Cell_type==ct) %>%
      select(fraction, {prams}) %>% 
      cor()
    out = cor_mat[1,] %>% as.data.frame()
    names(out)=ct
    return(out)
  })
  all <- all[-c(1), ]
  return(all)
}


i=1
for(i in 1:6){
  print(i)
  object <- readRDS(meta$spata2_new[i])
  nodes <- object@spatial[[1]]$graph$nodes
  edges <- object@spatial[[1]]$graph$edges
  
  all_drugs <- as.character(na.omit(nodes$drug %>% unique()))
  
  all_drugs_list = map(all_drugs, .f=function(drug_select){
    prams <- names(nodes)[c(18:24, 27:46)]
  cell_types <- names(cc)

  all = getCellCor(object, nodes, cell_types, k=4, drug_select, prams)

  all_sub <- 
    all %>% t() %>% 
    as.data.frame() %>% 
    rownames_to_column("annotation_level_4_new") %>% 
    left_join(., colors) %>% 
    select(-annotation_level_4_new) %>% 
    select(annotation_level_2, {prams}) %>% 
    group_by(annotation_level_2) %>% 
    summarise_all(.funs = mean) %>% 
    as.data.frame()
  })
  names(all_drugs_list) <- all_drugs
  
  saveRDS(all_drugs_list, paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))
  
  
}



matrix_list1 <- map(1:c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$nodrug 
  rownames(all_sub) <- all_sub$annotation_level_2
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })
matrix_list2 <- map(1:c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$`8mMK`
  rownames(all_sub) <- all_sub$annotation_level_2
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })
matrix_list3 <- map(1:c(1:2, 5:6), .f=function(i){
  all_sub <- readRDS(paste0(root_out, "/", meta$samples[i], "_cell_cor.RDS"))$`5uMgabazine`
  rownames(all_sub) <- all_sub$annotation_level_2
  #all_sub <- all_sub[,-c("firing_rate_whole_recording")]
  all_sub <- all_sub[,-c(1,6)]
  return(all_sub)
  })

matrix_list <- matrix_list3
all_sub <- Reduce("+", matrix_list) / length(matrix_list)
all_sub <- all_sub[colors$annotation_level_2 %>% unique(), ]
mat = all_sub[c(1:5, 7,9,10,16,17),] %>% t() %>% na.omit()

pheatmap::pheatmap(mat,
                   cluster_rows = F, cluster_cols = F,
                   color = colorRampPalette(RColorBrewer::brewer.pal(9, "PiYG") )(50))


```


# WGCNA
## Prepare WGCNA
```{r}
library(FNN)
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
getSumCounts <- function(object, nodes, sample_names, k=2, drug_select, prams){

  cells <- 
    object %>% 
    joinWithFeatures(features=cell_types)
  
  # subset drug
  n_inp = 
    nodes[!is.na(nodes$drug), ] %>% 
    group_by(id) %>% 
    summarise(x=mean(x), y=mean(y))

  
  ## KNN
  knn_result <- get.knnx(data = data.frame(x=cells$x, y=cells$y), query = n_inp[,c("x", "y")], k = 5)
  
  

  ## Summarize
  NN = map_dfr(1:nrow(knn_result$nn.index), .f=function(i){
      
    cell_mat = cells[knn_result$nn.index[i,], "barcodes"] 
    cell_mat <- 
      cell_mat %>% 
      mutate(
        dist = mean(knn_result$nn.dist[i,]),
        id = n_inp$id[i])
      
    return(cell_mat)
})
  
  NN <- left_join(NN, n_inp %>% select(id, x, y))
  NN <- left_join(NN, data.frame(barcodes = cells$barcodes, xend = cells$x, yend=cells$y))
  
  #ggplot()+
  #  geom_point(data = data.frame(x=cells$x, y=cells$y), mapping=aes(x,y))+
  #  geom_point(data = n_inp[,c("x", "y")],mapping=aes(x,y), color="red")+
  #  geom_segment(data = NN, mapping=aes(x=x,y=y,xend = xend,yend = yend))
  
  ## Get the Transcriptis
  counts = map_dfc(unique(NN$id), .f=function(i){
    
    counts = SPATA2::getCountMatrix(object)[,NN %>% filter(id==i) %>% pull(barcodes)] %>% as.matrix() %>% rowSums() %>% as.data.frame()
    names(counts)[1]=paste0(sample_names, "_", i)
    return(counts)
  }, .progress=T)
  
  
  
  
  
  return(counts)
}


i=1
mea_transcription = map(1:nrow(meta), .f=function(i){
  print(i)
  object <- readRDS(meta$spata2_new[i])
  nodes <- object@spatial[[1]]$graph$nodes
  sample_names = meta$samples[i]
  counts = 
    getSumCounts(object, nodes, k=5, sample_names=sample_names) %>% 
    as.matrix()
  return(counts)
})

## Create and integrate Seurat data:
print(i)
object <- readRDS(meta$spata2_new[i])
nodes <- object@spatial[[1]]$graph$nodes
sample_names = meta$samples[i]
n_inp = 
    nodes[!is.na(nodes$drug), ] %>% 
    filter(.$drug == "5uMgabazine") %>% 
    group_by(id,) %>% 
    select(id, {prams}) %>% 
    summarise_all(.funs = mean)

seurat <- Seurat::CreateSeuratObject(mea_transcription[[1]])

## Add Node features
seurat$sample = sample_names
seurat$mea_id = colnames(seurat) %>% str_remove(., pattern=sample_names) %>%  str_remove(., pattern="_")
seurat@meta.data <- cbind(seurat@meta.data, left_join(data.frame(id = seurat$mea_id), n_inp))

seurat_all = seurat

for(i in c(2,5,6)){
  print(i)
object <- readRDS(meta$spata2_new[i])
nodes <- object@spatial[[1]]$graph$nodes
sample_names = meta$samples[i]
n_inp = 
    nodes[!is.na(nodes$drug), ] %>% 
    filter(.$drug == "5uMgabazine") %>% 
    group_by(id,) %>% 
    select(id, {prams}) %>% 
    summarise_all(.funs = mean)

seurat <- Seurat::CreateSeuratObject(mea_transcription[[i]])

## Add Node features
seurat$sample = sample_names
seurat$mea_id = colnames(seurat) %>% str_remove(., pattern=sample_names) %>%  str_remove(., pattern="_")
seurat@meta.data <- cbind(seurat@meta.data, left_join(data.frame(id = seurat$mea_id), n_inp))


seurat_all <- merge(seurat_all, seurat)

}

library(Seurat)
seurat_all <- JoinLayers(seurat_all)

saveRDS(seurat_all, paste0(root_out,"/seurat_all_MEA.RDS"))

seurat_all <- readRDS(paste0(root_out,"/seurat_all_MEA.RDS"))
```


## WGCNA:

```{r}
library(WGCNA)
library(hdWGCNA)
library(igraph)
library(patchwork)

seurat_obj = readRDS(paste0(root_out,"/seurat_all_MEA.RDS"))

## Run Preprocessing
seurat_obj <- 
  seurat_obj %>% 
  FindVariableFeatures() %>% 
  NormalizeData() %>% 
  ScaleData() %>% 
  RunPCA()
## Run WGCNA
seurat_obj <- 
  hdWGCNA::SetupForWGCNA(seurat_obj, gene_select = "fraction",fraction = 0.01,wgcna_name = "vis") %>% 
  hdWGCNA::SetDatExpr(assay="RNA") %>% 
  hdWGCNA::TestSoftPowers()
wrap_plots(hdWGCNA::PlotSoftPowers(seurat_obj), ncol=2)
seurat_obj <- hdWGCNA::ConstructNetwork(seurat_obj,
                                        soft_power=8,
                                        deepSplit=2,
                                        minModuleSize=20,
                                        tom_name='MEA_Analysis',
                                        overwrite_tom=TRUE)
hdWGCNA::PlotDendrogram(seurat_obj, main='WGCNA dendrogram')
seurat_obj <- hdWGCNA::ModuleEigengenes(seurat_obj, assay = "RNA",npcs = 10)
seurat_obj <- hdWGCNA::ModuleConnectivity(seurat_obj)


```


```{r}
options(future.globals.maxSize= 8500*1024^2)

seurat_obj <- RunModuleUMAP(
  seurat_obj,
  n_hubs = 5,
  n_neighbors=5,
  min_dist=0.1,
  spread=2,
  wgcna_name = 'vis',
  target_weight=0.05,
  supervised=TRUE
)

p=ModuleUMAPPlot(
  seurat_obj,
  edge.alpha=0.25,
  sample_edges=TRUE,
  edge_prop=0.2, 
  label_hubs=2,
  return_graph=T,
  keep_grey_edges=FALSE)

```

```{r}

library(igraph)
library(scattermore)
library(ggrepel)
seurat_obj@misc$vis$MEs[ ] %>% pheatmap::pheatmap( cluster_rows = T)
plotdf <- seurat_obj@misc$vis$module_umap %>% filter(color!="grey60")
plotdf2 <- as_long_data_frame(p)
plotdf2 <- plotdf2 %>% filter(from_name %in% plotdf$gene)

# Create Plot of Expression Network
plot_p <- ggplot()+theme_void()+coord_fixed()
dat1 <- plotdf2 %>% filter(from_kME>0.5)
plot_p <- plot_p+geom_segment(data=dat1, mapping = aes(x=from_UMAP1, y=from_UMAP2, xend=to_UMAP1, yend=to_UMAP2), size=dat1$from_kME*0.1, alpha=dat1$from_kME*0.1)
plot_p <- plot_p+geom_scattermore(data=plotdf, mapping=aes(x=UMAP1, y=UMAP2),size=plotdf$kME*2, alpha=0.5, color=gplots::col2hex(plotdf$color), pointsize=4)
plot_p <- plot_p+theme_void()+geom_text_repel(data=plotdf %>% group_by(module) %>% summarise(UMAP1=mean(UMAP1), UMAP2=mean(UMAP2)), mapping=aes(x=UMAP1, y=UMAP2,label=module))
plot_p <- plot_p+geom_text_repel(data=plotdf %>% group_by(module) %>% top_n(., 15, wt=kME), mapping=aes(x=UMAP1, y=UMAP2,label=gene),size=2)
plot_p+
     theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.25),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1), 
          axis.text.y = element_text(colour="black"))
```

```{r}
## DE Analysis
ModuleCorrelogram(seurat_obj)
```

## WGCNA Correlation analysis
```{r}

seurat_obj <- ModuleTraitCorrelation(
  seurat_obj,
  traits = prams)


seurat_obj <- ModuleTraitCorrelation(
  seurat_obj,
  traits = c("betweenness_centrality", "degree","degree_centrality", "firing_rate", "nonrandom_spikes" ))


mt_cor <- GetModuleTraitCorrelation(seurat_obj)

col=colorRampPalette(rev(c(RColorBrewer::brewer.pal(9, "RdBu"))))


k = t(mt_cor$cor$all_cells) %>% as.data.frame() %>% pheatmap::pheatmap()

k$tree_row$order

mat <- t(mt_cor$cor$all_cells)
p_mat = t(mt_cor$pval$all_cells)


library(ggcorrplot)
ggcorrplot(mat[k$tree_row$order, k$tree_col$order] %>% as.data.frame(), 
           method = "circle", 
           outline.color = "black", 
           lab_size=1,
           insig="pch",
           pch="X",
           p.mat=p_mat[k$tree_row$order, k$tree_col$order])+
  scale_fill_gradientn(colours = col(50), name="", limits = c(-0.4,0.4), oob=scales::squish) +
  guides(fill = guide_colourbar(barwidth = 0.3, barheight = 8, ticks =F, frame.colour="black"), label=F)+
  ggtitle("Module correlation to treatment condition")+
  xlab("")+ylab("")+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.25),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1), 
          axis.text.y = element_text(colour="black"))

```

## Get gene set enrichment

```{r}
library(DOSE)
library(enrichplot)
library(clusterProfiler)

module = "saddlebrown"

GSEA_all <- map(unique(seurat_obj@misc$vis$wgcna_modules$module), .f=function(module){
  out_kME <- 
  data.frame(SYMBOL = seurat_obj@misc$vis$wgcna_modules$gene_name,
           kME = seurat_obj@misc$vis$wgcna_modules[,paste0("kME_", module)])
out_genes <- bitr(out_kME$SYMBOL, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
out_kME <- left_join(out_kME,out_genes)
  out_kME <- out_kME %>% filter(!is.na(ENTREZID)) %>% arrange(desc(kME))
values = out_kME$kME;names(values) <- out_kME$ENTREZID
GSEA <- clusterProfiler::gseKEGG(values)

return(GSEA)
})


names(GSEA_all) <- unique(seurat_obj@misc$vis$wgcna_modules$module)

GSEA_GO <- map(unique(seurat_obj@misc$vis$wgcna_modules$module), .f=function(module){
  out_kME <- 
  data.frame(SYMBOL = seurat_obj@misc$vis$wgcna_modules$gene_name,
           kME = seurat_obj@misc$vis$wgcna_modules[,paste0("kME_", module)])
out_genes <- bitr(out_kME$SYMBOL, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db")
out_kME <- left_join(out_kME,out_genes)
  out_kME <- out_kME %>% filter(!is.na(ENTREZID)) %>% arrange(desc(kME))
values = out_kME$kME;names(values) <- out_kME$ENTREZID
GSEA <- clusterProfiler::gseGO(values,ont="ALL", keyType = "ENTREZID",OrgDb = "org.Hs.eg.db")

return(GSEA)
})
names(GSEA_GO) <- unique(seurat_obj@misc$vis$wgcna_modules$module)

enrichplot::dotplot(GSEA_GO$royalblue, showCategory=60)

names(GSEA_all)
module = "royalblue"
GSEA_GO[[module]]@result %>% 
  arrange(desc(enrichmentScore)) %>% 
  head(20) %>%
  arrange((enrichmentScore)) %>% 
  select(Description, enrichmentScore, p.adjust) %>% 
  mutate(Description = factor(Description, levels=Description)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=enrichmentScore, y=Description, size=-log(p.adjust), fill=enrichmentScore), shape=21)+
  scale_fill_gradientn(colors=viridis::inferno(50))+
  guides(fill = guide_colourbar(barwidth = 0.3, barheight = 8, ticks =F, frame.colour="black"), label=F)+
  xlab("")+ylab("")+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.25),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1), 
          axis.text.y = element_text(colour="black"))



seurat_obj <- Seurat::AddModuleScore(seurat_obj, features=list(wnk_spak_genes=wnk_spak_genes), name="pw")


inp_MEs = seurat_obj@misc$vis$MEs
pw=seurat_obj$pw1

map(1:ncol(inp_MEs), ~cor(pw, inp_MEs[,.x])) %>% unlist() %>% as.data.frame() %>% mutate(modules=names(inp_MEs)) %>% arrange(desc(.))

royalblue

ggplot(mapping=aes(x=inp_MEs[,"royalblue"], y=pw))+geom_point()+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.25),
          axis.text.x = element_text(colour="black",
                                     angle = 0, vjust = 0.5, hjust=1), 
          axis.text.y = element_text(colour="black"))


## Module to GBMap 

scRef <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBMap_Neuron.RDS")
scRef$annotation_level_2[is.na(ref$annotation_level_2)] <- "Neurons"


scRef$annotation_level_4 <- ref$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")

colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4_new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(., "-", "_") %>% 
  str_replace_all(., " ", "_") %>% 
  str_replace_all(., "/", "_")
```


## Create a module score:
```{r}

# Create module Score
#scRef <- ref
#scRef <- Seurat::AddModuleScore(scRef, features = list_genes, name=names(list_genes), assay = "SCT")
#names(list_genes)

modules <- seurat_obj@misc$vis$wgcna_modules$color %>% unique()
modules <- modules[2:length(modules)]
list_genes_imp <- map(.x=modules[1:length(modules)], .f=function(i){

  importance <- 
    seurat_obj@misc$vis$wgcna_modules %>% 
    filter(module==i) %>% 
    select(paste0("kME_",i))

  return(importance)
  
})
names(list_genes_imp) <- modules

# get the amount of genes from each signature that is expressed

Module_signature <- map(list_genes_imp, .f=function(genes){
  
  # weighted mean expression
  mat <- Seurat::GetAssayData(scRef, assay = "SCT")
  mat <- mat[na.omit(match(rownames(genes), rownames(mat))),] %>% as.data.frame()
  
  mat <- mat*genes[rownames(mat), ]
  
  mean <- colMeans(mat)
  out <- apply(mat, 2, function(x){ length(which(x!=0))/length(x) })*100
  names(out) <- colnames(mat)
  mean=mean*out

  
  
  return(list(mean, out))
  
  
  
  
}, .progress = T)
names(Module_signature) <- paste0(names(list_genes_imp), 1:length(names(list_genes_imp)))

for(n in names(Module_signature)){
  scRef@meta.data[,n] <- Module_signature[[n]][[1]]
}

names(Module_signature)

```

## Plot modules 
```{r}

module <- "lightgreen28"

library(Seurat)
f <- FeaturePlot(scRef, features = module, order=T)

library(scattermore)
p <- ggplot(data=f$data)+
  geom_scattermore(mapping=aes(UMAP_1,UMAP_2), pointsize = 8, color="black")+
  geom_scattermore(mapping=aes(UMAP_1,UMAP_2), pointsize = 7, color="white")+
  geom_scattermore(mapping=aes(UMAP_1,UMAP_2, color=!!sym(module)), pointsize=2)

col=colorRampPalette(c("#FFFFFF", RColorBrewer::brewer.pal(9, "Reds")))
p <- p+
    scale_colour_gradientn(colours = col(50),
                           #limit=c(0,1.5), 
                           oob = scales::squish, na.value = "white")+
    ylab("UMAP2")+xlab("UMAP1")+
    theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"), 
          axis.text.y = element_text(colour="black"))+
    coord_fixed()
p



```


```{r}
Module_score <- names(Module_signature)
data_enrichment <- scRef@meta.data[,c("annotation_level_4", Module_score)]

enrichment <- 
  data_enrichment %>% 
  group_by(annotation_level_4) %>% 
  summarise_all(.funs = base::mean)

size <- map_dfc(1:length(Module_signature), ~ as.data.frame(Module_signature[[.x]][[2]]))
names(size) <- names(Module_signature)
size <- 
  size %>% 
  mutate(annotation_level_4=scRef$annotation_level_4) %>% 
  group_by(annotation_level_4) %>% 
  summarise_all(.funs = mean)


#tumors <- colors[colors$annotation_level_2 %in% c("Differentiated-like", "Stem-like"), ]$annotation_level_4
tumors <- colors$annotation_level_4_new
enrichment <- enrichment %>% filter(annotation_level_4 %in% tumors) %>% as.data.frame()
rownames(enrichment) <- enrichment$annotation_level_4
enrichment$annotation_level_4 <- NULL

size <- size %>% filter(annotation_level_4 %in% tumors) %>% as.data.frame()
rownames(size) <- size$annotation_level_4
size$annotation_level_4 <- NULL

enrichment_mat <- scales::rescale(as.matrix(enrichment), c(0,1)) %>% as.data.frame()
size <- scales::rescale(as.matrix(size), c(0,1)) %>% as.data.frame()
dim(enrichment)
dim(size)

enrichment <- 
  enrichment_mat %>%
  rownames_to_column(var = "gene") %>%
  gather(cell, count, -gene)

size <- 
  size %>%
  rownames_to_column(var = "gene") %>%
  gather(cell, count, -gene)

enrichment$size <- size$count

colors <- 
  colors %>% 
  filter(annotation_level_4!="Neuron")



col=colorRampPalette(c("#FFFFFF", RColorBrewer::brewer.pal(9, "Reds")))
#col=colorRampPalette(rev(RColorBrewer::brewer.pal(9, "PRGn")))
pheatmap::pheatmap(enrichment_mat %>% cor())


#rank <- map_dbl(1:ncol(seurat_obj@misc$vis$MEs), ~ cor(seurat_obj@misc$vis$MEs[.x], seurat_obj$random_spikes))
#names(rank) <- colnames(seurat_obj@misc$vis$MEs)
#rank <- rank[modules]
#names(rank) <- paste0(names(rank), 1:length(names(rank)))
#rank <- rank[order(rank)]

# "pink24", "royalblue14",

#names(Module_signature)
enrichment <- 
  enrichment %>% 
  filter(cell %in%c("skyblue37", "lightcyan31", "lightgreen28", "darkred4", "darkolivegreen26"))

#names(Module_signature)
#enrichment <- 
#  enrichment %>% 
#  filter(cell %in%c("royalblue14", "blue9"))

library(ggcorrplot)
plot_1 <- 
  #ggcorrplot(enrichment, method = "circle", outline.color = "black")+
  ggplot(enrichment)+
  geom_point(aes(x = gene, y = cell, size = size, fill=count),colour="black",pch=21)+
  scale_y_discrete(limits = names(rank))+
  scale_x_discrete(limits = colors$annotation_level_4_new)+
  scale_fill_gradientn(colours = col(50), name="", limits=c(0,0.001), oob=scales::squish)+
  scale_size(range = c(0,3))+
  guides(fill = guide_colourbar(barwidth = 0.3, barheight = 8, ticks =F, frame.colour="black"), label=F)+
  ggtitle("Expression vs Metabolism")+
  xlab("")+ylab("")+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black",
                                     angle = 90, vjust = 0.5, hjust=1), 
          axis.text.y = element_text(colour="black"))+
  coord_fixed()

plot_1



```









```{r}

modules <- c("tan", "salmon", "yellow", "red", "purple")

gcSample <- map(.x=modules, .f=function(i){
  genes <- 
    seurat_obj@misc$vis$wgcna_modules %>% 
    filter(module==i) %>% 
    rownames()
  
  genes <- c(unique(genes))
})

all <- unlist(gcSample)
names(gcSample) <- modules


library(DOSE)
library(enrichplot)
library(clusterProfiler)
xx <- compareCluster(gcSample, 
                     fun="enrichGO",
                     ont ="BP", 
                     keyType = "SYMBOL", 
                     OrgDb = "org.Hs.eg.db")

edo <- pairwise_termsim(xx)

plot <- 
  emapplot(edo, showCategory=50, cex_category=0.01, 
           pie.params = list(pie = "equal"), 
           cex_line = 0.3,
           layout="graphopt", min_edge=0.15, repel=T)

res_map <- 
  edo@compareClusterResult %>% 
  filter(Description %in% plot$data$name) %>% 
  group_by(Description, Cluster) %>% 
  summarise(p = mean(p.adjust)) %>% 
  ungroup() 

anno <- data.frame(name=unique(res_map$Description), type=map(.x=unique(res_map$Description), .f=function(i){
  c <- res_map %>% filter(Description==i) %>% pull(Cluster)
  return(c[res_map %>% filter(Description==i) %>% pull(p) %>% which.min()] %>% as.character())
  }) %>% unlist())

library(ggrepel)
plot$data <- plot$data %>% left_join(., anno, by="name")

col <- plot$data$type %>% unique() %>% rev() %>% gplots::col2hex()

ggplot(data=plot$data)+
  plot$layers[[1]]+
  geom_point(mapping = aes(x=x, y=y, fill=type, size=size), colour="black",pch=21)+
  scale_fill_manual(values=col)+
  theme_bw() +
  xlab("Dim 1")+ylab("Dim 2")+
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"), 
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  geom_text_repel(mapping= aes(x,y, label=name), size=2)+
  ggtitle("GeneSet Enrichment Analysis")




```







```{r}
object_list <- readRDS(samples_deconv_df %>% filter(Samples_ID=="404_R") %>% pull(path_deconv))

for(i in 1:length(object_list$cell_map)){object_list$cell_map[[i]]$match <- as.character(object_list$cell_map[[i]]$match)}
cell_types <- object_list$cell_map %>% map_dfr(., ~.x %>% as.data.frame())

## Annotation with GBMap_new
intersect = intersect(cell_types$match %>% unique(), colnames(ds.ref2))
new_annotation_spatial = ds.ref2@meta.data[intersect,] %>% as.data.frame() %>% rownames_to_column("match") %>% select(match, predicted.cells)
cell_types <- left_join(cell_types, new_annotation_spatial)
cell_types[!is.na(cell_types$predicted.cells), "celltypes"] <- 
  cell_types[!is.na(cell_types$predicted.cells), "predicted.cells"] 

## plot seurat
ds.ref2$cells = factor(ds.ref2$predicted.cells, levels = myeloid_color$cell_type)
color_sc_myeloid = myeloid_color$color;names(color_sc_myeloid)=myeloid_color$cell_type
DimPlot( subset(ds.ref2, cells = intersect) , 
         group.by="cells", reduction = "ref.umap", 
         label = F, raster=T, pt.size = 20, 
         raster.dpi = c(1512, 1512))+
  scale_color_manual(values=color_sc_myeloid)



table(cell_types$celltypes)

object <- object_list$object %>% SPATA2::updateSpataObject()
colors <- readRDS("<USER_HOME>/Desktop/RecurrentGBM/SingleCellRef/GBM_Neuron_colors.RDS")
colors$annotation_level_4new <- 
  colors$annotation_level_4 %>% 
  str_replace_all(" ", "_") %>% 
  str_replace_all("-", "_") %>% 
  str_replace_all("/", "_")
# Add new myeloid lables 
cc_myeloid = colors[colors$annotation_level_2=="Myeloid", "colors"] %>% pull(colors)

colors <- rbind(colors %>% filter(annotation_level_2!="Myeloid"),
                data.frame(annotation_level_2="Myeloid", 
           annotation_level_4 = new_annotation_spatial$predicted.cells %>% unique(), 
           n=1, colors = colorRampPalette(c(cc_myeloid[3], cc_myeloid[11]))(length(new_annotation_spatial$predicted.cells %>% unique())),
           annotation_level_4new = new_annotation_spatial$predicted.cells %>% unique()))

cell_types$celltypes <- factor(cell_types$celltypes, levels = colors$annotation_level_4new)
cc <- colors$colors;names(cc) <- colors$annotation_level_4new

cell_types

ggplot(cell_types, aes(x, y, fill = celltypes))+
  geom_point(shape=21, size=1)+
  scale_fill_manual(values=cc)+
  theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")

## Plot only microglia cells 
myeloid_color
cc <- rep("lightgrey",nrow(colors)) ; names(cc) <- colors$annotation_level_4new
cc[c(myeloid_color %>% filter(cell_type %in% names(cc)) %>% pull(cell_type))] = 
  myeloid_color %>% filter(cell_type %in% names(cc)) %>% pull(color)

library(ggforce)
ggplot(cell_types, aes(x, y, group = -1L)) +
    geom_voronoi_tile(aes(fill = celltypes), max.radius = 10, colour = 'black', linewidth=0.1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  SPATA2::ggpLayerAxesSI(object)+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")


```


# Graph Visulatisation of the Electrogenomic data:
```{r}
#Colors:
cc <- colors$colors;names(cc) <- colors$annotation_level_4_new
i=6
object <- readRDS(meta$spata2_new[i])
plotSurface(object)

object@used_genesets = rbind(data.frame(ont = "wnk_spak_genes", gene=wnk_spak_genes), object@used_genesets)


limits=c(0.1, 0.8)
p = plotSurface(object, color_by = "wnk_spak_genes", alpha_by="wnk_spak_genes", display_image=F, limits=limits, oob=scales::squish)+
  scale_alpha(limits=limits,oob=scales::squish)
p

cell_types <- object@spatial[[1]]$cytospace


p = plotSurface(object, pt_alpha=0)

library(ggforce)
p=ggplot()+
  geom_voronoi_tile(data =cell_types, mapping=aes(x=x_sc,y=y_sc, group = -1L, fill = CellType), 
                      max.radius = 3, colour = 'black', linewidth=0.1, alpha=1)+
    scale_fill_manual(values=cc)+
    theme_bw() +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"),
          axis.text.y = element_text(colour="black"))+
    coord_fixed()+
  theme(legend.text = element_text(size=6),
        legend.key.size = unit(0.3, "cm"),
        legend.title = element_text(size=8))+
  xlab("Dimension x-space [cm]")+
  ylab("Dimension y-space [cm]")+
  Seurat::NoLegend()


## Add MEA Grid
library(ggnewscale)
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges

## Create graph:
library(igraph)       # graph object
library(GephiForR)    # ForceAtlas 2
library(ggraph)       # modern plotting

f_nodes = nodes %>% filter(drug=="8mMK")
f_edges = edges %>% filter(drug=="8mMK")

g = graph_from_data_frame(d=edges %>% filter(drug=="8mMK") %>% select(from, to, weight))

# run ForceAtlas 2
coords_out <- layout.forceatlas2(
  g,
  iterations   = 400,   
  linlog       = TRUE, 
  gravity      = 1,
  scalingratio = 10,
  plotlabels=F
)

coords_out <- coords_out %>% as.data.frame() %>% rownames_to_column("id")
f_nodes <- left_join(f_nodes, coords_out, by="id")
f_edges <- cbind(f_edges, 
                 cbind(
                   left_join(data.frame(id=f_edges$from), coords_out) %>% select(V1, V2) %>% rename(x_coords_from = V1, y_coords_from = V2),
                   left_join(data.frame(id=f_edges$to), coords_out) %>% select(V1, V2) %>% rename(x_coords_to = V1, y_coords_to = V2)
                   )
                 )


# draw it
ggplot()+
  geom_point(f_nodes, mapping=aes(x=V1, y=V2, size=degree))+
  geom_segment(data = f_edges %>% arrange(weight), 
               mapping=aes(x=x_coords_from,y=y_coords_from,xend=x_coords_to,yend=y_coords_to, color =weight, 
                           alpha=weight, 
                           size=weight), alpha=0.1)+
  scale_size(range=c(0.1,0.3))+
  scale_color_viridis_c(, 
                        limits=c(0, 400), 
                        oob=scales::squish)
  



cond <- nodes$drug %>% unique()
drug_select = c(nodes$drug %>% unique())[3]
weight_lim = 0.4

n_inp = nodes[!is.na(nodes$drug), ] %>% filter(.$drug == drug_select)
e_inp = edges %>% filter(.$drug==drug_select) %>% filter(weight>weight_lim)


p+
  ggnewscale::new_scale_color()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=1)+
  geom_segment(data = e_inp %>% arrange(weight), 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight), alpha=0.1)+ 
  scale_size(range=c(0.1,0.3))+
  scale_color_viridis_c(, 
                        limits=c(0, 400), 
                        oob=scales::squish)

ggplot()+
  geom_point(data = n_inp, mapping=aes(x=x,y=y,alpha=degree), size=2, color="black")+
  geom_segment(data = e_inp, 
               mapping=aes(x=x,y=y,xend=xend,yend=yend, color =weight, alpha=weight, size=weight))+ 
  scale_size(range=c(0.1,1))+
  scale_color_viridis_c(option="B", limits=c(0.3, 0.7), oob=scales::squish)


```

## 3D Visualisation two layer:

```{r}
## Graph visualisation in 3D : 

# ── packages ────────────────────────────────────────────────────────────
library(dplyr)      # pipes & joins
library(plotly)     # 3-D plotting
library(igraph)     # your graph object  (optional if you already have edgelist)

# ── 1. build one tidy "node table" --------------------------------------
force_coords   <- data.frame(id = coords_out$id, x_layout = coords_out$V1, y_layout = coords_out$V2)
spatial_coords <- data.frame(id = f_nodes$id, x_spat   = f_nodes$x, y_spat   = f_nodes$y)

## Scale Coords into the same size:

spatial_coords$x_spat <- scales::rescale(spatial_coords$x_spat, range(force_coords$x_layout) )
spatial_coords$y_spat <- scales::rescale(spatial_coords$y_spat, range(force_coords$y_layout) )

nodes <- force_coords %>% inner_join(spatial_coords, by = "id")
nodes <- left_join(nodes, f_nodes %>% select(id, degree, region))

# ── 2. choose the vertical gap between the layers -----------------------
z_layout <- 0      # ForceAtlas layer
z_spat   <- 50     # spatial layer (any numeric value)

# ── 3. convert igraph edges to a data frame (or use your own) -----------
edges <- f_edges %>% filter(weight>20)

# ── 3. helper that returns a list of plotly traces ---------
library(viridisLite)   # nice perceptual palette
library(scales)        # col_numeric()

weight_pal <- col_numeric(
  viridis(256, option = "D"),           # or any palette you like
  domain = range(edges$weight, na.rm = TRUE)
)


make_edge_traces_layout <- function(edge_tbl, xs, ys, z) {
  edge_tbl %>%
    left_join(nodes %>% select(id, x0 = {{ xs }}, y0 = {{ ys }}),
              by = c("from" = "id")) %>%
    left_join(nodes %>% select(id, x1 = {{ xs }}, y1 = {{ ys }}),
              by = c("to"   = "id")) %>%
    mutate(col = weight_pal(weight)) %>%           # map colour here
    pmap(function(x0, y0, x1, y1, col, ...) {
      list(
        x = c(x0, x1), y = c(y0, y1), z = c(z, z),
        type  = "scatter3d", mode = "lines",
        line  = list(width = 2, color = col),      # <- coloured by weight
        showlegend = FALSE
      )
    })
}

# ── 4. create the three trace groups -----------------------
z_layout <- 50
z_spat   <- 0   # change to taste

edge_traces_layout <- make_edge_traces_layout(edges, x_layout, y_layout, z_layout)
edge_traces_spat   <- make_edge_traces_layout(edges, x_spat,   y_spat,   z_spat)


library(dplyr)
library(purrr)
library(RColorBrewer)   # Set2 palette

region_levels <- sort(unique(nodes$region))
pal <- setNames(
  brewer.pal(max(3, length(region_levels)), "Set2")[seq_along(region_levels)],
  region_levels
)

connector_traces <- nodes %>%
  pmap(function(id, x_layout, y_layout, x_spat, y_spat, region, ...) {
    list(
      x = c(x_layout, x_spat),
      y = c(y_layout, y_spat),
      z = c(z_layout,  z_spat),
      type  = "scatter3d",
      mode  = "lines",
      line  = list(width = 2,
                   color = pal[[region]]),   # <- use the palette here
      hovertext = id,                        # optional: tool-tip
      showlegend = FALSE
    )
  })

# ── 5. build the final plotly figure -----------------------
p <- plot_ly() %>%

  # ForceAtlas-2 nodes
  add_trace(
    data = nodes,
    x = ~x_layout, 
    y = ~y_layout, 
    z = z_layout,
    size = ~degree,
    type = "scatter3d", mode = "markers",
    marker = list(color = "tomato"),
    name   = "ForceAtlas2 layout"
  ) %>%

  # spatial-coordinate nodes
  add_trace(
    data = nodes,
    x = ~x_spat, 
    y = ~y_spat, 
    z = z_spat,
    color = ~region,
    colors = "Set2",
    type = "scatter3d", mode = "markers",
    marker = list(size = 8),
    name   = "Spatial layer"
  )

# add edges and connectors
#all_traces <- c(edge_traces_layout, edge_traces_spat, connector_traces)
all_traces <- c(edge_traces_layout, connector_traces)
for (tr in all_traces) {
  p <- exec(add_trace, p, !!!tr)   # splice ‘tr’ into add_trace()
}

p <- p %>%
  layout(
    scene = list(
      xaxis = list(
        title          = "",      # no axis title
        showticklabels = FALSE,   # <- hides the numbers & tick text
        showgrid       = TRUE,   # optional: kill the grey grid
        zeroline       = TRUE    # optional: hide the axis line itself
      ),
      yaxis = list(
        title          = "",
        showticklabels = FALSE,
        showgrid       = TRUE,
        zeroline       = TRUE
      ),
      zaxis = list(
        title          = "",
        showticklabels = FALSE,
        showgrid       = TRUE,
        zeroline       = TRUE
      )
    )
  )



p %>%
  layout(
    scene = list(
      xaxis = list(title = ""),
      yaxis = list(title = ""),
      zaxis = list(title = "")
    )
  )


```




## 3D Visualisation three layer:

```{r}
i=6
object <- readRDS(meta$spata2_new[i])
plotSurface(object)
object@used_genesets = rbind(data.frame(ont = "wnk_spak_genes", gene=wnk_spak_genes), object@used_genesets)

plotSurface(object, color_by="wnk_spak_genes", smooth=F)


## Add MEA Grid
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges
## Filter condition:
f_nodes = nodes %>% filter(drug=="8mMK")
f_edges = edges %>% filter(drug=="8mMK")


## Test 
plotSurface(object, color_by="wnk_spak_genes", smooth=F, pt_alpha=0)

plotSurface(object, color_by="wnk_spak_genes", smooth=F, display_image=T)+
  ggnewscale::new_scale_color()+
  geom_point(data = f_nodes, mapping = aes(x=x, y=y, fill=region),alpha=0.5, size=4,shape=21, color="white")


## Create Layout 
g = graph_from_data_frame(d=f_edges %>% select(from, to, weight))
coords_out <- layout.forceatlas2(
  g,
  iterations   = 300,   
  linlog       = TRUE, 
  gravity      = 10,
  scalingratio = 10,
  plotlabels=F
)

```


```{r}
## Graph visualisation in 3D : 
coords_out <- coords_out %>% as.data.frame() %>% rownames_to_column("id")
f_nodes <- left_join(f_nodes, coords_out, by="id")
f_edges <- cbind(f_edges, 
                 cbind(
                   left_join(data.frame(id=f_edges$from), coords_out) %>% select(V1, V2) %>% rename(x_coords_from = V1, y_coords_from = V2),
                   left_join(data.frame(id=f_edges$to), coords_out) %>% select(V1, V2) %>% rename(x_coords_to = V1, y_coords_to = V2)
                   )
                 )

## Plot Inputs
force_coords   <- data.frame(id = coords_out$id, x_layout = coords_out$V1, y_layout = coords_out$V2)
spatial_coords <- data.frame(id = f_nodes$id, x_spat   = f_nodes$x, y_spat   = f_nodes$y)
expression <- joinWithGeneSets(object, gene_sets = "wnk_spak_genes") %>% select(x,y,wnk_spak_genes)
names(expression)[3] <- "value"

## Test layout
ggplot()+
  geom_point(data = force_coords, mapping=aes(x=x_layout,y=y_layout), size=2, color="black")+
  geom_segment(data = f_edges %>% filter(weight>20), 
               mapping=aes(x=x_coords_from,y=y_coords_from,xend=x_coords_to,yend=y_coords_to, color =weight, alpha=weight, size=weight))+ 
  scale_size(range=c(0.1,1))+
  scale_color_viridis_c(option="B", 
                        #limits=c(0.3, 0.7), 
                        oob=scales::squish)



## Re-Scale layout coords

force_coords$x_layout <- scales::rescale(force_coords$x_layout, range(spatial_coords$x_spat) )
force_coords$y_layout <- scales::rescale(force_coords$y_layout, range(spatial_coords$y_spat) )
nodes <- force_coords %>% inner_join(spatial_coords, by = "id")
nodes <- left_join(nodes, f_nodes %>% select(id, degree, region))


# ── 2. choose the vertical gap between the layers -----------------------
z_layout <- 10      # ForceAtlas layer
z_spat   <- 5     # spatial layer (any numeric value)
z_exp   <- 0

# ── 3. convert igraph edges to a data frame (or use your own) -----------
edges <- f_edges %>% filter(weight>20)

# ── 3. helper that returns a list of plotly traces ---------
library(viridisLite)   # nice perceptual palette
library(scales)        # col_numeric()

weight_pal <- col_numeric(
  viridis(256, option = "D"),           # or any palette you like
  domain = range(edges$weight, na.rm = TRUE)
)


make_edge_traces_layout <- function(edge_tbl, xs, ys, z) {
  edge_tbl %>%
    left_join(nodes %>% select(id, x0 = {{ xs }}, y0 = {{ ys }}),
              by = c("from" = "id")) %>%
    left_join(nodes %>% select(id, x1 = {{ xs }}, y1 = {{ ys }}),
              by = c("to"   = "id")) %>%
    mutate(col = weight_pal(weight)) %>%           # map colour here
    pmap(function(x0, y0, x1, y1, col, ...) {
      list(
        x = c(x0, x1), y = c(y0, y1), z = c(z, z),
        type  = "scatter3d", mode = "lines",
        line  = list(width = 2, color = col),      # <- coloured by weight
        showlegend = FALSE
      )
    })
}

edge_traces_layout <- make_edge_traces_layout(edges, x_layout, y_layout, z_layout)
edge_traces_spat   <- make_edge_traces_layout(edges, x_spat,   y_spat,   z_spat)


library(dplyr)
library(purrr)
library(RColorBrewer)   # Set2 palette

region_levels <- sort(unique(nodes$region))
pal <- setNames(
  brewer.pal(max(3, length(region_levels)), "Set2")[seq_along(region_levels)],
  region_levels
)

connector_traces <- nodes %>%
  pmap(function(id, x_layout, y_layout, x_spat, y_spat, region, ...) {
    list(
      x = c(x_layout, x_spat),
      y = c(y_layout, y_spat),
      z = c(z_layout,  z_spat),
      type  = "scatter3d",
      mode  = "lines",
      line  = list(width = 2,
                   color = pal[[region]]),   # <- use the palette here
      hovertext = id,                        # optional: tool-tip
      showlegend = FALSE
    )
  })

# ── 5. build the final plotly figure -----------------------
p <- plot_ly() %>%

  # ForceAtlas-2 nodes
  add_trace(
    data = nodes,
    x = ~x_layout, 
    y = ~y_layout, 
    z = z_layout,
    size = ~degree,
    type = "scatter3d", mode = "markers",
    marker = list(color = "tomato"),
    name   = "ForceAtlas2 layout"
  ) %>%

  # spatial-coordinate nodes
  add_trace(
    data = nodes,
    x = ~x_spat, 
    y = ~y_spat, 
    z = z_spat,
    color = ~region,
    colors = "Set2",
    type = "scatter3d", mode = "markers",
    marker = list(size = 4),
    name   = "Spatial layer"
  ) %>% 
  # Expression layer
  add_trace(
    data = expression,
    x = ~x, 
    y = ~y, 
    z = z_exp,
    color = ~value,
    colors = "inferno",
    type = "scatter3d", mode = "markers",
    marker = list(size = 4),
    name   = "Expression layer"
  )

# add edges and connectors
#all_traces <- c(edge_traces_layout, edge_traces_spat, connector_traces)
all_traces <- c(edge_traces_layout)#, connector_traces)
for (tr in all_traces) {
  p <- exec(add_trace, p, !!!tr)   # splice ‘tr’ into add_trace()
}

p <- p %>%
  layout(
    scene = list(
      xaxis = list(
        title          = "",      # no axis title
        showticklabels = FALSE,   # <- hides the numbers & tick text
        showgrid       = TRUE,   # optional: kill the grey grid
        zeroline       = TRUE    # optional: hide the axis line itself
      ),
      yaxis = list(
        title          = "",
        showticklabels = FALSE,
        showgrid       = TRUE,
        zeroline       = TRUE
      ),
      zaxis = list(
        title          = "",
        showticklabels = FALSE,
        showgrid       = TRUE,
        zeroline       = TRUE
      )
    )
  )



p %>%
  layout(
    scene = list(
      xaxis = list(title = ""),
      yaxis = list(title = ""),
      zaxis = list(title = "")
    )
  )


```


```{r}
## Add MEA Grid
nodes <- object@spatial[[1]]$graph$nodes
edges <- object@spatial[[1]]$graph$edges

unique(nodes$drug)


## Filter condition:
f_nodes = nodes %>% filter(drug=="5uMgabazine")
f_edges = edges %>% filter(drug=="5uMgabazine")


## Test 
plotSurface(object, color_by="wnk_spak_genes", smooth=F, pt_alpha=0)

plotSurface(object, color_by="wnk_spak_genes", smooth=F, display_image=T)+
  ggnewscale::new_scale_color()+
  geom_point(data = f_nodes, mapping = aes(x=x, y=y, fill=region),alpha=0.5, size=4,shape=21, color="white")


## Create Layout 
g = graph_from_data_frame(d=f_edges %>% select(from, to, weight))
coords_out <- layout.forceatlas2(
  g,
  iterations   = 300,   
  linlog       = TRUE, 
  gravity      = 5,
  scalingratio = 10,
  plotlabels=F
)

## Graph visualisation in 3D : 
coords_out <- coords_out %>% as.data.frame() %>% rownames_to_column("id")
f_nodes <- left_join(f_nodes, coords_out, by="id")
f_edges <- cbind(f_edges, 
                 cbind(
                   left_join(data.frame(id=f_edges$from), coords_out) %>% select(V1, V2) %>% rename(x_coords_from = V1, y_coords_from = V2),
                   left_join(data.frame(id=f_edges$to), coords_out) %>% select(V1, V2) %>% rename(x_coords_to = V1, y_coords_to = V2)
                   )
                 )

## Plot Inputs
force_coords   <- data.frame(id = coords_out$id, x_layout = coords_out$V1, y_layout = coords_out$V2)
spatial_coords <- data.frame(id = f_nodes$id, x_spat   = f_nodes$x, y_spat   = f_nodes$y)
expression <- joinWithGeneSets(object, gene_sets = "wnk_spak_genes") %>% select(x,y,wnk_spak_genes)
names(expression)[3] <- "value"

## Test layout
ggplot()+
  geom_point(data = f_nodes, mapping=aes(x=V1,y=V2, fill=region), size=2,shape=21, color="black")+
  geom_segment(data = f_edges %>% filter(weight>10), 
               mapping=aes(x=x_coords_from,y=y_coords_from,xend=x_coords_to,yend=y_coords_to, color =weight, alpha=weight, size=weight))+ 
  scale_size(range=c(0.1,1))+
  scale_color_viridis_c(option="B", 
                        #limits=c(0.3, 0.7), 
                        oob=scales::squish)+
  theme_bw() +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_rect(colour = "black", size=0.5),
          axis.text.x = element_text(colour="black"), 
          axis.text.y = element_text(colour="black"))


```


